### 2020-05-12 manuscript.Rmd
create manuscript file with vancouver csl and bibl

---
output:
  output: pdf_document
bibliography: bibliography.bib
csl: vancouver.csl
---

# Multiple deprivation and geographic distance to community sport events — achieving equitable access to parkrun in England

  Schneider PP$^{*,1}$, Smith RA$^{1}$, Bullas AM$^{2}$, Quirk H$^{2}$, Bayley T$^{1}$, Haake SSJ$^{2}$, Brennan A$^{1}$, Goyder E$^{1}$

*$^{1}$School of Health and Related Research, University of Sheffield, Sheffield, UK.; $^{2}$Advanced Wellbeing Research Centre, Sheffield Hallam University, Sheffield, UK.*

## Introduction

Insufficient physical activity is a leading cause of disease and disability worldwide.[@who2009] For the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic background are disproportionately affected by physical inactivity-related diseases. Increasing the physical activity levels of the population is therefore high on the public health agenda: it not only has the potential to improve quality of life, reduce mortality rates and alleviate the strain on health and social care services, but also reduce the gap in health inequalities.

However, designing effective public health interventions that increase population physical activity is considerable challenge.[@ref] Implementing such interventions in a way that does not increase health inequalities might even be more difficult, as studies have shown that programmes to increase physical activity often fail to reach deprived communities and those most in need.[@white2009; @rideout2018]

parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professional with respect to both. Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2004; @wiltshire2018]

Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely driven by *demand*, not *need*. It might therefore be the case that parkrun events are primarily located in areas that are relatively well-off anyway, while people living in more deprived communities may not have the same opportunities to participate. In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@sportengland2018]

The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England is equitable across areas with different levels of deprivation. Secondly, to identify 200 optimal locations for future events, in order to improve geographic access for deprived communities.

## Methods

This study is an ecological analysis of the geographic and socio-economic disparities in participation in parkrun events in England. The observational period covered 49 weeks in 2018 (1st January to 10th December). All analyses were conducted on the level of Lower Layer Super Output Area (LSOA). LSOAs are census areas with a population of approximately 1,700, which divide England into 32,844 geographic units.

\bigskip

We collected and combined location (longitude and latitude) and attribute data of LSOAs with information on parkrun events and public green spaces. An overview of the relevant measures and the data sources is provided in Table 1.

\bigskip

In this study, we included all 455 English parkrun events which were in operation during the observational period. Events held in prisons (n=7) were excluded. Data on 143,822 public green spaces were retrieved from an open dataset of Ordnance Survey.[9] In the absence of more detailed information (e.g. terrain), we considered all public parks, gardens, and playing fields in England with an area of 0.1 km$^2$ (e.g. 316m x 316m) or more potentially suitable for hosting parkrun events (n=2,842). We set this lower bound in order to improve the specificity of our estimates and avoid identifying very small community parks. At the same time we wanted to ensure that parks such as mileend park (which has a parkrun and is 0.014 km$^2$) would be identified in our analysis.

<!--# %%% TABLE 1 %%%
# \begin{figure}[!t]
#     \centering
#     \includegraphics[width=1\textwidth]{tbl1.jpg}
#     \label{fig:tbl1}
# \end{figure}-->

\bigskip

\subsection*{Analysis}
Mean, standard deviation, median, interquartile range, and range were used as descriptive statistics. The associations between IMD, distance to the nearest parkrun event, and parkrun participation on the LSOA level were investigated using generalised linear models with a Poisson distribution and a log link. The dependent variable was the total number of runs between 1st January and 10th December 2018. To model the outcome as a rate (runs per 1,000 population per week), the log of the person weeks was used as an offset variable.

\bigskip

We first fitted bivariate models to study the effect of IMD and distance on participation separately. The correlation between both predictors was assessed using Pearson’s correlation coefficient. Subsequently, we used a multivariable model to investigate the independent effects of IMD and distance on participation. The Akaike (AIC) and the Bayesian Information Criterion (BIC), as well as the McFadden’s pseudo R$^{2}$ were used to assess model fit.

\bigskip

\subsection*{Identifying optimal locations for new parkrun events}

\subsubsection*{Rationale and objective}

We conducted a location-allocation analysis to solve the following problem: parkrun UK received funding to start 200 additional parkrun events. There are 2,842 public green spaces in England in which new events could be set up. Which 200 locations should be selected?

\bigskip

parkrun’s stated aim of ‘increasing participation from lower socio-economic groups’[8] was operationalised in the following way: \textit{find the set of 200 green spaces, which maximise the number of additional runs at parkrun events, weighted by LSOAs’ squared IMD scores}. By using the squared IMD scores as weights, the generation of new runs from more deprived areas is prioritised over new runs from less deprived areas.
We also explored the following alternative specifications: 1. minimise deprivation-weighted distances; 2. maximise the total (i.e. unweighted) number of runs; 3. Minimise (unweighted) distances.

\subsubsection*{Location-allocation analysis}
We developed a flexible algorithm to select the optimal new parkrun event locations from the set of candidate green spaces. The algorithm consists of two steps: firstly, for each green space, we evaluate how many new runs (weighted by squared LSOA IMD score) would be generated by setting up a new parkrun event at the respective location. Secondly, the green space with the greatest effect on new runs is selected. This procedure is repeated 200 times.

\bigskip

More formally, we define that for any candidate green space location j, the objective function $f(j|E)$ provides the sum of parkrun runs $r_{i}$ over all LSOA $i$, weighted by the squared IMD score $w_{i}^{2}$, given the set of established parkrun event locations $E=\{e_{1},e_{2},...,e_{455}\}$:

\[ f(j | E)=\sum_{i=1}^{32844}{ w_{i}^{2} * r_{i j}} \]

In the absence of causal estimates, we use the Poisson regression model specified above to predict the expected number of runs $r_{ij}$ for LSOA $i$ based on its IMD score $w_{i}$, its (linear) distance to the nearest parkrun event $d_{ij}$, and its population $p_{i}$. The functional form is given below.

\[ E(r_{ij} |w_{i},d_{ij},p_{i})=\exp{( \beta_{0}+\beta_{1}*w_{i}+\beta_{2}*d_{ij}+\ln(p_{i})+\in}) \]

Filling-in the parameter coefficients (see table 3), we derive the following formula:
\[ \hat{r_{ij}} = \exp{}(-5.402-0.048*w_{i}-0.082*d_{ij}+\ln(p_{i}))\]

Note that $j$ can have an effect on $r_{ij}$ through $d_{ij}$: setting up a new event at location $j$ will reduce the distance to the nearest event for some LSOA $i$. This means, we evaluate the distances from LSOA $i$’s location $l_{i}$ to all established parkrun event locations $\{e_{1},e_{2},...,e_{455}\} \in E$, denoted $\overline{l_{i} e_{1}}, \overline{l_{i} e_{2}}, \ldots, \overline{l_{i} e_{455}}$, and to the candidate green space location $j$, denoted $\overline{l_{i}j}$, and then take the minimum value, i.e. $d_{ij} = \min(\overline{l_{i}j}, \overline{l_{i} e_{1} }, \overline{l_{i} e_{2}}, \ldots, \overline{l_{i} e_{455}})$.

\bigskip

The expected change in the objective function is computed for all candidate locations j in the set of the available green spaces $C=\{c_{1},c_{2}, \ldots, c_{2842}\}$, and the location with the maximum value is selected. The selection function is expressed in the following formula:

\[\underset{j \in \mathcal{C}}{\arg \max } f(j | E)\]

In order to identify the optimal set of 200 new locations $O=\{o_{1},o_{2}, \ldots, o_{200}\}$, the selection procedure is repeated 200 times. At each step $k$, the best candidate green space location $o_{k} \in C$ is selected, added to the set of established parkrun events $E$, and removed from the set of available green spaces $C$. Thereby, the $k$th new parkrun event location is taken into account when selecting the $k$th+1 location. The full algorithm can be described in pseudo code in the following way:

\bigskip

\subsection*{Interactive map, source code and data}

We mapped all relevant data points, including LSOAs, green spaces, parkrun events, and recommended new event locations, and created an interactive map, which can be accessed online [16]. An annotated version of the R source code and all data that were used to generate the results of this study are provided on a repository under a creative commons license [17].

\bigskip

\subsection*{Ethical approval}
Ethical approval was obtained from the Sheffield Hallam University Ethics Committee (ER10776545). We did not collect any personal information, but only used aggregate secondary data. The parkrun Research Board approved this research project, and four of its members (AMB, HQ, EG, SSJH) were actively involved in the intepretation of findings and writing of this manuscript.

## Results

## Discussion

## Conclusion

## References

<div id="refs"></div>

## Appendix

### 2020-05-12 src/src4_analysis.R

@@ -15,9 +15,6 @@
     length(event_sp$course) # n parkrun events
     summary.stats(event_sp$srvd_pop) # served population
     summary.stats(event_sp$srvd_lsoa) # served LSOA
-    summary.stats( event_sp$Mn_prtc) # mean participants per event
-    summary.stats( event_sp$Mn_vlnt) # mean volunteers
-    summary.stats(event_sp$Age_yrs) # age of events 
     
     # pop living_within_1km 
     round(sum(lsoa_sp$pop[lsoa_sp$mn_dstn<=1])/sum(lsoa_sp$pop),4)*100

### 2020-05-12 src/src2_data_proc.R
rm participation data

@@ -51,11 +51,11 @@
     # parkrun event data 
       # Parkrun UK. 2019. 
       # Retrieved from: http://www.parkrun.org.uk/
-      event_sp = read.csv("./input/parkrun_data/events_corrected.csv", stringsAsFactors = F)
+      event_sp = read.csv("./input/parkrun_data/event_info_20181231.csv", stringsAsFactors = F)
     
     # LSOA parkrun participation data (raw data from parkrunUK, filtered and processed)
       # Parkrun UK. 2018. 
-      runs_per_lsoa = read.csv("./input/parkrun_data/runs_per_lsoa_010118_101218.csv", stringsAsFactors = F)
+      # runs_per_lsoa = read.csv("./input/parkrun_data/runs_per_lsoa_010118_101218.csv", stringsAsFactors = F)
     
 
       
@@ -81,8 +81,8 @@
       lsoa_pop$pop_km2 = as.numeric(gsub(",","",lsoa_pop$pop_km2))
       lsoa_sp = merge(lsoa_sp,lsoa_pop,by=c("code"),sort=F)
     
-    # add time var (observational period = 49 weeks)
-      lsoa_sp@data$week = 49
+    # # add time var (observational period = 49 weeks)
+    #   lsoa_sp@data$week = 49
     
     # prepare imd scores 
       lsoa_imd = lsoa_imd %>% 
@@ -138,15 +138,15 @@
       event_sp = merge(event_sp,srvd_pop,by="course")
     
   
-  # PARTICIPATION data
-    # prepare runs per week, per 1,000 population 
-      lsoa_sp = merge(lsoa_sp,runs_per_lsoa,by="code",sort=F)
-      lsoa_sp$run_count[is.na(lsoa_sp$run_count)] = 0
-      lsoa_sp$runs_pmil.week = ((lsoa_sp$run_count/49)/lsoa_sp$pop)*1000 # 49 weeks observational period
-      
+  # # PARTICIPATION data
+  #   # prepare runs per week, per 1,000 population
+  #     lsoa_sp = merge(lsoa_sp,runs_per_lsoa,by="code",sort=F)
+  #     lsoa_sp$run_count[is.na(lsoa_sp$run_count)] = 0
+  #     lsoa_sp$runs_pmil.week = ((lsoa_sp$run_count/49)/lsoa_sp$pop)*1000 # 49 weeks observational period
+    
 # CLEAN UP
[…]

### 2020-05-14 manuscript.Rmd
full analysis revision

@@
 
-This study is an ecological analysis of the geographic and socio-economic disparities in participation in parkrun events in England.
-The observational period covered 49 weeks in 2018 (1st January to 10th December).
-All analyses were conducted on the level of Lower Layer Super Output Area (LSOA).
-LSOAs are census areas with a population of approximately 1,700, which divide England into 32,844 geographic units.
+### Study design
 
-\bigskip
+  This study is a cross-sectional ecological analysis of the socio-economic disparities in geographic access to parkrun events in England at the end of 2018.
+All analyses were conducted on the level of Lower Layer Super Output Areas (LSOAs), which divide England into 32,844 geographic units with, on average, have a population of approximately 1,700.
+We assessed the relationship between access, defined as the distance (as the crow flies) to the nearest parkrun event, and socio-economic deprivation, measured using the Index of Multiple Deprivation.
+In addition, we used information on green spaces in England to conducted a simple location-allocation analysis, in order to identify 200 locations for future parkrun events, that maximise access for access for the population.
 
-We collected and combined location (longitude and latitude) and attribute data of LSOAs with information on parkrun events and public green spaces.
-An overview of the relevant measures and the data sources is provided in Table 1.
+### Data sources
 
-\bigskip
+  For this study, we combined data on three types of geospatial entities: 1) LSOAs, 2) parkrun event locations, and 3) public green spaces.
 
-In this study, we included all 455 English parkrun events which were in operation during the observational period.
-Events held in prisons (n=7) were excluded.
-Data on 143,822 public green spaces were retrieved from an open dataset of Ordnance Survey.[9] In the absence of more detailed information (e.g.
-terrain), we considered all public parks, gardens, and playing fields in England with an area of 0.1 km$^2$ (e.g.
-316m x 316m) or more potentially suitable for hosting parkrun events (n=2,842).
-We set this lower bound in order to improve the specificity of our estimates and avoid identifying very small community parks.
-At the same time we wanted to ensure that parks such as mileend park (which has a parkrun and is 0.014 km$^2$) would be identified in our analysis.
+  1) The geographic locations of all 32,844 LSOAs, defined by the coordinates of its population-weighted centroid, 2017 total population estimates, and the 2015 Index of Multiple Deprivation (IMD) were retrieved from the Office for National Statistics [@ons2018; @ref; @ref].
 
-<!--# %%% TABLE 1 %%%
-# \begin{figure}[!t]
-#     \centering
-#     \includegraphics[width=1\textwidth]{tbl1.jpg}
-#     \label{fig:tbl1}
-# \end{figure}-->
+  2) We included all 465 public parkrun events which were in operation in England by December 12th 2018 - on this date, Sport England announced that their plan to provide funding for setting up 200 additional parkrun events across England.[@ref] The coordinates of all 465 events were manually scrapped from the parkrun UK website.[@ref]
 
-\bigskip
+  3) The location coordinates of 143,822 public green spaces were retrieved from an open dataset of Ordnance Survey.[9] parkrun events are held in various settings and terrains, and not always require a 5km loop - some events have courses that involve running back and forth.
+After evaluating existing parkrun event courses, we decided to consider all public parks, gardens, and playing fields in England with an area of 0.1 km$^2$ (e.g.
+316m x 316m) or more potentially suitable for hosting events (n=2,842).
 
-\subsection*{Analysis}
+### Variables
+
+  Access to parkrun was measured in terms of the distance (as the crow flies) to the nearest event.
+For each of the 32,844 LSOAs, we computed the geodesic distances between its population-weighted centroid and all 465 parkun events and then selected the shortest distance (in km).
+
+  The socio-economic deprivation of LSOAs was measured using IMD.
+It is a measure of measure of relative deprivation, which combines 37 indicators from seven domains (income, employment, education and skills, health and disability, crime, housing and services, and living environment) into a single score.
+The score ranges from 0 (least deprived) to 100 (most deprived).[@smith2015]
+
+  Further covariates (e.g.
+population density or demographics), which are likely to affect the availability of parkrun events, were not taken into account, because we did not aim to assess to what extent deprivation independently 'explains' access.
+Rather, we sought to evaluate whether or not people living in deprived areas actually have better or worse access, given all other circumstances.
+
+### Analysis
+
   Mean, standard deviation, median, interquartile range, and range were used as descriptive statistics.
-The associations between IMD, distance to the nearest parkrun event, and parkrun participation on the LSOA level were investigated using generalised linear models with a Poisson distribution and a log link.
-The dependent variable was the total number of runs between 1st January and 10th December 2018.
-To model the outcome as a rate (runs per 1,000 population per week), the log of the person weeks was used as an offset variable.
+The correlation between IMD and the distance to the nearest parkrun event on the LSOA level were investigated using the Pearson and Spearman method.
+The correlation coefficients were computed using the LSOAs' total population as weights.
+A scatter plot with a fitted linear regression model are presented to illustrate the relationship.
+Our hypothesis was that more deprived areas had worse access, i.e.
+longer distances to the nearest parkrun event than more affluent areas.
 
-\bigskip
-
-We first fitted bivariate models to study the effect of IMD and distance on participation separately.
-The correlation between both predictors was assessed using Pearson’s correlation coefficient.
-Subsequently, we used a multivariable model to investigate the independent effects of IMD and distance on participation.
-The Akaike (AIC) and the Bayesian Information Criterion (BIC), as well as the McFadden’s pseudo R$^{2}$ were used to assess model fit.
-
-\bigskip
-
-\subsection*{Identifying optimal locations for new parkrun events}
-
-\subsubsection*{Rationale and objective}
+#### Identifying optimal locations for new parkrun events
 
@@
 There are 2,842 public green spaces in England in which new events could be set up.
-Which 200 locations should be selected?
+Which 200 locations should be selected, in order to maximise access for the greatest number of people?
 
-\bigskip
-
-parkrun’s stated aim of ‘increasing participation from lower socio-economic groups’[8] was operationalised in the following way: \textit{find the set of 200 green spaces, which maximise the number of additional runs at parkrun events, weighted by LSOAs’ squared IMD scores}.
-By using the squared IMD scores as weights, the generation of new runs from more deprived areas is prioritised over new runs from less deprived areas.
-We also explored the following alternative specifications: 1. minimise deprivation-weighted distances; 2. maximise the total (i.e. unweighted) number of runs; 3. Minimise (unweighted) distances.
-
-\subsubsection*{Location-allocation analysis}
-We developed a flexible algorithm to select the optimal new parkrun event locations from the set of candidate green spaces.
-The algorithm consists of two steps: firstly, for each green space, we evaluate how many new runs (weighted by squared LSOA IMD score) would be generated by setting up a new parkrun event at the respective location.
-Secondly, the green space with the greatest effect on new runs is selected.
+  More specifically, the objective was to minimise the population-weighted total sum of distances between all LSOAs and their nearest parkrun event.
+To identify the optimal 200 green spaces, we applied a simple greedy algorithm that consisted of two steps.
+Firstly, for each green space, we evaluate how setting up a parkrun event would affect the sum of distances, given the locations of all existing events (i.e.
+for how many LSOAs would this green space then be the nearest parkrun event, and by how much would it decrease the respective distances).
+Secondly, the green space with the greatest effect is selected and added to the set of existing parkrun events.
 This procedure is repeated 200 times.
@@
 
-More formally, we define that for any candidate green space location j, the objective function $f(j|E)$ provides the sum of parkrun runs $r_{i}$ over all LSOA $i$, weighted by the squared IMD score $w_{i}^{2}$, given the set of established parkrun event locations $E=\{e_{1},e_{2},...,e_{455}\}$:
+  More formally, the first step of the algorithm is defined by the following equation:
 
-\[ f(j | E)=\sum_{i=1}^{32844}{ w_{i}^{2} * r_{i j}} \]
+$$\underset{c\in C}{\operatorname{arg min}}\sum_{i=1}^{32,844}{ d_i(E \cap c) * p_{i}}$$
 
-In the absence of causal estimates, we use the Poisson regression model specified above to predict the expected number of runs $r_{ij}$ for LSOA $i$ based on its IMD score $w_{i}$, its (linear) distance to the nearest parkrun event $d_{ij}$, and its population $p_{i}$.
-The functional form is given below.
+  The function yields the candidate green space $c$, from the set of all 2,842 green spaces $C$, which minimises the sum of the products of the minimum distances and the total population of all LSOAs, whereby $p_i$ denotes the population of LSOA $i$, and $d_i(E \cap c)$ denotes the distance to the nearest parkrun event, which can either be an existing event from the set $E = \{e_{1},e_{2},...,e_{465}\}$, or the candidate green space $c$, whichever is nearest.
 
-\[ E(r_{ij} |w_{i},d_{ij},p_{i})=\exp{( \beta_{0}+\beta_{1}*w_{i}+\beta_{2}*d_{ij}+\ln(p_{i})+\in}) \]
+  In order to identify the optimal 200 new locations for setting up new parkun events, the selection procedure is repeated 200 times.
+At each step, the single best candidate green space location is selected, added to the set of established parkrun events $E$, and removed from the set of available green spaces $C$.
+This means, the optimal green space selected at step $k$ is is taken into account when selecting the $k$th+1 location.
 
-Filling-in the parameter coefficients (see table 3), we derive the following formula:
-\[ \hat{r_{ij}} = \exp{}(-5.402-0.048*w_{i}-0.082*d_{ij}+\ln(p_{i}))\]
+#### Evaluating the location allocation analysis
 
-Note that $j$ can have an effect on $r_{ij}$ through $d_{ij}$: setting up a new event at location $j$ will reduce the distance to the nearest event for some LSOA $i$.
-This means, we evaluate the distances from LSOA $i$’s location $l_{i}$ to all established parkrun event locations $\{e_{1},e_{2},...,e_{455}\} \in E$, denoted $\overline{l_{i} e_{1}}, \overline{l_{i} e_{2}}, \ldots, \overline{l_{i} e_{455}}$, and to the candidate green space location $j$, denoted $\overline{l_{i}j}$, and then take the minimum value, i.e.
-$d_{ij} = \min(\overline{l_{i}j}, \overline{l_{i} e_{1} }, \overline{l_{i} e_{2}}, \ldots, \overline{l_{i} e_{455}})$.
+  We assessed the overall impact of setting up 200 new parkrun events at the identified optimal location on the geographic access to parkrun events in England.
+We also conducted a distributional effectiveness analysis, in which we grouped LSOAs into IMD quintiles (most, more, median, less, least deprived) and subsequently assessed how setting up the 200 new events would affect access in each of those subgroups.
 
-\bigskip
+### Data and source code availability
 
-The expected change in the objective function is computed for all candidate locations j in the set of the available green spaces $C=\{c_{1},c_{2}, \ldots, c_{2842}\}$, and the location with the maximum value is selected.
-The selection function is expressed in the following formula:
+  All data and the R source code that were used to generate the results of this study are provided on an open repository.[@ref]
 
-\[\underset{j \in \mathcal{C}}{\arg \max } f(j | E)\]
+### Ethical approval
 
-In order to identify the optimal set of 200 new locations $O=\{o_{1},o_{2}, \ldots, o_{200}\}$, the selection procedure is repeated 200 times.
-At each step $k$, the best candidate green space location $o_{k} \in C$ is selected, added to the set of established parkrun events $E$, and removed from the set of available green spaces $C$.
-Thereby, the $k$th new parkrun event location is taken into account when selecting the $k$th+1 location.
-The full algorithm can be described in pseudo code in the following way:
-
-\bigskip
-
-\subsection*{Interactive map, source code and data}
-
-We mapped all relevant data points, including LSOAs, green spaces, parkrun events, and recommended new event locations, and created an interactive map, which can be accessed online [16].
-An annotated version of the R source code and all data that were used to generate the results of this study are provided on a repository under a creative commons license [17].
-
-\bigskip
-
-\subsection*{Ethical approval}
   Ethical approval was obtained from the Sheffield Hallam University Ethics Committee (ER10776545).
 We did not collect any personal information, but only used aggregate secondary data.
-The parkrun Research Board approved this research project, and four of its members (AMB, HQ, EG, SSJH) were actively involved in the intepretation of findings and writing of this manuscript.
+The parkrun Research Board approved this research project, and four of its members (AMB, HQ, EG, SSJH) were actively involved in the interpretation of findings and writing of this manuscript.
 
@@
 
+### Descriptive statistics
+
+  As of 12th December 2018, approximately 7%, 69%, and 91% of the English population lived within 1, 5, and 10km of a parkrun event.
+Only 578,043 people (1% of the English population) lived more than 20km from an event.
+The average (SD) and median (IQR) distance to the nearest parkrun event were 4.65 (4.22) and 3.39 (1.99-5.83).
+The worst access was observed for 2,259 people living on the Isles of Scilly, who live about 76.44km (as the crow flies) away from the next parkrun event on the mainland.
+Further descriptive statistics are provided in table 1.
+
+| **Table 1: Descriptive statistics of LSOAs and parkrun events** |
+| ---|---|---|--- |
+| <td colspan=4>**LSOA (n =n 32,844) ** |
+| ---|---|---|--- |
+| |Mean (SD) | Median (Q1-Q3) | Range |
+| Population | 1,693.44 (405.35) | 1,612 (1,452-1,834) | 362-13,404 |
+| IMD score | 21.67 (15.59) | 17.40 (9.65-30.07) | 0.48-92.60 |
+| Distance to the nearest parkrun event location | 4.65 (4.22) | 3.39 (1.99-5.83) | 0.04-76.44 |
+|
+| <td colspan=4>**Parkrun Events (n = 465)** |
+| ---|---|---|--- |
+| Served LSOAs\* | 70.63 (42.81) | 62 (40-87) | 6-350 |
+| Served population\* | 119,612 (74,290) | 103,952 (68,837-151,488) | 7,855-628,010 |
+| ---|---|---|--- |
+ | \* Number of LSOAs or population for which a given parkrun event is the nearest. |
+
+### Association between Deprivation and Access
+
+There was a negative relationship between IMD and the distance to the nearest parkrun event: the (population weighted) Pearson and Spearman correlation coefficients were -0.15 and -0.18, indicating a small effect.
+This means, more deprived LSOAs tended to have shorter distances to the nearest parkun event, i.e.
+better access.
+
+The distributional analysis of distances by IMD quantile showed that people living in the 20% most deprived LSOAs  had the best geographic access, with an average and median distance to the nearest parkrun event of 3.51 and 2.79 km.
+Depending on the metric, the worst access was observed for LSOAs in the middle (mean distance = 3.36km) or the less deprived group (median distance = 3.93km).
+Further results of the distributional analysis are shown in Table 2 in the following section.
+
+### Optimal locations for new parkrun events
+
+Figure 1 shows the parkrun events (black circles) that existed on 12th December 2018 alongside recommendations for 200 additional event locations (red triangles), which maximise overall access to parkrun for the greatest number of people.
+The numbers correspond to the rank, where 1 is the location which would improve access the most.
+The names and exact locations of the selected 200 green spaces are provided in the appendix.
+
+![](./output/map_static.jpeg)
+**Figure 1: Map of England showing current parkrun events (blue circles) and recommended new event locations (red triangles) ranked in descending order of estimated effect on overall population-weighted access.
+Information on all 200 identified optimal green space locations are provided in the appendix.**
+
+We estimated that setting up new parkrun events in those 200 green spaces would improve access for around 16.5 million people (30% of th population) from 9,854 LSOAs.
+For those, the distance to the nearest event would, on average, be reduced by 4.09 km (SD = 3.97).
+The percentage of people who live within 5km of a parkrun would increase from 69% to 81.9%.
+
+Table 2 shows the difference in geographic access before and after the introduction of the 200 new parkrun events at the identified optimal green spaces.
+Results are presented for the overall effect and stratified by IMD quantiles.
+The 200 new events amplified the relationship between IMD and geographic access further: the population-weighted Pearson and Spearman correlation coefficients changed from -0.15 and -0.18 before, to -0.20 and -0.23 afterwards.
+Nevertheless, the improvement in access was smallest for LSOAs in the most deprived quantile.
+
+Figure 2 shows the distribution of the nearest distances between LSOAs and parkrun events before and after the introduction of the 200 new optimal event locations.
+
+![](./output/hist.png)
+**Figure 2: caption**
+
+```{r}
+# Table 2
+# quantile(lsoa_sp$imd_sc,seq(0,1,0.2))
+# cbind(mndst_tbl,mndst_new_tbl)
+```
+
 ## Discussion
+
+<!-- Main findings -->
+- Access to parkrun is good
+- Contrary to our expectation, we did not find that access is better for people living in less deprived areas. On the contrary, people in the most deprived areas had the best geographic access.
+- setting up 200 new events in optimal green spaces would further improve access for a large part of the population. 
+- Setting up new events in the way we proposed would again benefit
+
+<!-- further considerations -->
+- an explanation for the negative association between IMD and distance is urban/rural: deprived areas tend to cluster in cities and this is also where most parkruns take place
+
+<!-- link to relevant studies -->
+
+<!-- strengths & limitations -->
+- first quantitative analysis
+- geographic access is not travel distance or travel time
+- greenspace: not all are suitable, and some suitable places (promenades, beaces, national parks) are not included.
+However, green spaces are distrubuted over England and it seems unlikely that any important area was missed completely.
+
+<!-- main discussion points -->
+1 ) - proximity only provides the opportunity to participate in parkrun events
+  - in a previous study we found that after controlling for distance and other factors, people from deprived areas are less likely to participate in parkrun events.
+  - this means, giving access is neccessary but not sufficient to increase participation from deprivied communities.
+  - parkrun should complement their efforts with...
+
+2 ) - future research should build on this work and develop a model to assess the (cost-)effectiveness of setting up new events not only in terms of improved potential access, but actual participation.
+This requires estimating causal and marginal effects.
+More, longitudinal data and more sophisticated modelling are required.
+
+<!-- end on positive note -->
+
+- parkrun is hugely succesfull and has the potential to improve levels of physcial activity of the pop?
 

### 2020-05-14 src/src1_functions.R

@@ -3,228 +3,40 @@
 # FUNCTIONS
   
 # convenient function to install (if neccessary) and load required packages
-  install_n_load = function(package){
-      for(i in 1:length(package)){
-        if(eval(parse(text=paste("require(",package[i],")")))==0) {
-          install.packages(package)
-        }
-      }
-      return (eval(parse(text=paste("require(",package,")"))))
-      }
-  
+install_n_load = function(package){
+  for(i in 1:length(package)){
+    if(eval(parse(text=paste("require(",package[i],")")))==0) {
+      install.packages(package[i])
+    }
+  }
+  return (eval(parse(text=paste("require(",package[i],")"))))
+}
+
+
 # compute basic descriptive stats
   summary.stats = function(x,round.dec=2){
-    mean.d = mean(x)
-    sd.d = sd(x)
-    median.d = median(x)
-    q25.d = quantile(x,c(.25))
-    q75.d = quantile(x,c(.75))
-    min.d = min(x)
-    max.d = max(x)
     
-    res = data.frame(mean.d,
-               sd.d,
-               median.d,
-               q25.d,
-               q75.d,
-               min.d,
-               max.d
+    res = data.frame(mean.d = mean(x),
+                     sd.d = sd(x),
+                     median.d = median(x),
+                     q25.d = quantile(x,c(.25)),
+                     q75.d = quantile(x,c(.75)),
+                     min.d = min(x),
+                     max.d = max(x)
     )
    round(res,round.dec) 
   }
   
-# extract some basic glm model results
-  extract_glm_values = function(model = m_dist_imd_runs){
-    coefs = names(model$coefficients)
-    betas = as.numeric(model$coefficients)
-    ci.95 = confint.default(model)
-    exp.ci.95 = formatC(exp(ci.95),digits=3,format="f")
-    ci.95 = formatC(ci.95,digits=3,format="f")
-    ci.95 = ifelse(substr(ci.95,1,1)=="-",ci.95,paste(" ",ci.95,sep=""))
-    exp.betas = formatC(exp(betas),digits=3,format="f")
-    betas = formatC(betas,digits=3,format="f")
-    betas = ifelse(substr(betas,1,1)=="-",betas,paste(" ",betas,sep=""))
[…]

### 2020-05-14 src/src2_data_proc.R

@@ -2,33 +2,33 @@
 # LOAD, MERGE, AND COMPUTE THE REQUIRED DATA
 
 # LOAD FUNCTIONS + PACKAGES
-  # Load functions
-  #  source("./src/src1_functions.R") # required
-  
   # Install and load all required packages
     install_n_load(c("raster",
                      "sp",
+                     "sf",
+                     "mapproj",
+                     "scales",
                      "geosphere",
                      "rgeos",
+                     "reshape2",
                      "dplyr",
                      "ggplot2",
+                     "eq5d",
+                     "cowplot",
                      "ggrepel",
+                     "lwgeom",
                      "maptools",
-                     "ggmap")
+                     "ggmap",
+                     "wCorr")
                    )
   
 # LOAD RAW DATA
-    # LSOA polygon geospatial data 
-      # Contains public sector information licensed under the Open Government Licence v3.
-      # Office for National Statistics (2011). 2011 Census: boundary data (England and Wales) [data collection].
-      # Retrieved from: https://data.gov.uk/dataset/fa883558-22fb-4a1a-8529-cffdee47d500/lower-layer-super-output-area-lsoa-boundaries
-      lsoa_sp = raster::shapefile("./input/England_lsoa_2011_sgen_clipped/england_lsoa_2011_sgen_clipped")
     
     # LSOA population-weighted centroids
       # Contains public sector information licensed under the Open Government Licence v3.
       # Office for National Statistics (2011). 2011 Census: boundary data (England and Wales) [data collection].
       # Retrieved from: https://data.gov.uk/dataset/a40f54f7-b123-4185-952f-da90c56b0564/lower-layer-super-output-areas-december-2011-population-weighted-centroids
-      lsoa_cntrds = raster::shapefile("./input/England_lsoa_2011_centroids/england_lsoa_2011_centroids")
+      lsoa_sp = raster::shapefile("./input/England_lsoa_2011_centroids/england_lsoa_2011_centroids")
       
     # LSOA population + LSOA area sizes
[…]

### 2020-05-14 src/src3_localoalg.R

@@ -1,0 +1,159 @@
+
+######################
+## parkrun event Location Allocation
+## find optimal locations for future parkrun events
+##################
+
+    top_n  = 200   # how many new locations are needed?
+  
+  # setup
+    events.coord = coordinates(event_sp)
+    
+    lsoa.coord = coordinates(lsoa_sp)
+    lsoa.n = length(lsoa_sp$code)
+    lsoa.pop = lsoa_sp$pop
+    min_dist_lsoa_events = lsoa_sp$mn_dstn*1000
+    objective_state.append = sum(min_dist_lsoa_events*lsoa.pop)  
+    
+    candidates.coord = coordinates(greens_sp)
+    candidate_index = 1:length(greens_sp$id)
+    
+  # Compute all distances between LSOA and candidate green spaces 
+  # ! ! ! cave: high memory demands ! ! !
+    dist_lsoa_candidates = distm(lsoa.coord,candidates.coord)
+  
+  # data frame to store results 
+    res_df = data.frame(index=NA,objective=NA,change=rep(NA,times=top_n+1),n=NA,pop=NA,mean=NA)
+    # first row (step 0) shows baseline
+    res_df[1,] = c(0,objective_state.append,0,0,0,0)
+  
+  t1 = Sys.time()
+# GREEDY LOC ALO ALG
+  # OUTER LOOP
+  for(k in 1:top_n){ # repeat as many times as new locations are to be identified
+    
+    timecheck = 0 # inner loop progress index
+    
+    len = length(candidate_index) # n candidates remaining
+    
+    # INNER LOOP
+    for(i in 1:len){ # search through all candidate parks
+      
+      # show progress as % 
+      progress = round((i/len),2)*100
+      if(abs(progress - timecheck)>0.1){
+        timecheck = progress
+        cat("\r Outer loop: ",k, "   /",top_n,"; inner loop: %",progress, sep="")
+        #     flush.console()
+      }
+      
+      # which lsoa had a lower mn_dstn if candidate i were a parkrun?
+      reassign = dist_lsoa_candidates[,i] < min_dist_lsoa_events
+      if(sum(reassign)>0){ # if at least one LSOA improves...
+        # evaluate the sum of the new objective function
+        new_trans_min_dist_pop_combi = rep(NA,times=lsoa.n)
+        new_trans_min_dist_pop_combi[reassign] = dist_lsoa_candidates[reassign,i]
[…]

### 2020-05-14 src/src3_recommendations.R

@@ -1,82 +1,0 @@
-
-### OPTIMAL NEW PARKRUN LOCATIONS
-# Code to reproduce optimal new parkrun event recommendations
-# source("./src/src2_data_proc.R")  # required
-    
-    time1 = Sys.time() # keeps track of time
-    
-    # specify GLM to estimate the effect of new runs on LSOAs
-    m_dist_imd_runs = glm(run_count ~ imd_sc + mn_dstn,
-                          data = lsoa_sp@data,
-                          family = poisson(link="log"),
-                          offset = log(pop*week))
-    
-    # model for distance only
-    m_dist_dist = lm(mn_dstn+0 ~ -1 + mn_dstn ,data = lsoa_sp@data)
-    
-    # new data frame for predictions
-    imd_runs_df = lsoa_sp@data[,c("pop","imd_sc","week")]
-   
-# MAX TOTAL RUNS IMD WEIGHTED
-    opt_loc_runs_imd.prep = LocAloAlg(candidates = greens_sp,
-                                 event_coordinates.M = event_sp, 
-                                 lsoa_centr = cbind(lsoa_sp@data$centr_lng,
-                                                    lsoa_sp@data$centr_lat),
-                                 data = imd_runs_df,
-                                 model = m_dist_imd_runs,
-                                 weights = lsoa_sp@data$imd_sc^2, 
-                                 top_n = 200)
-    shapefile(opt_loc_runs_imd.prep,filename="./output/opt_loc_runs_imd.prep.shp",overwrite = T)
-    gc()
-    gc()
-    
-# MAX TOTAL RUNS
-    opt_loc_runs.prep = LocAloAlg(candidates = greens_sp,
-                             event_coordinates.M = event_sp, 
-                             lsoa_centr = cbind(lsoa_sp@data$centr_lng,
-                                                lsoa_sp@data$centr_lat),
-                             data = imd_runs_df,
-                             model = m_dist_imd_runs,
-                             weights = 1, 
-                             top_n = 200,
-                             maximise = T)
-    
-    shapefile(opt_loc_runs.prep,filename="./output/opt_loc_runs.prep.shp",overwrite = T)
-    gc()
-    gc()
-    
-# MIN DISTANCE IMD WEIGHTED
-    opt_loc_dist_imd.prep = LocAloAlg(candidates = greens_sp,
-                                 event_coordinates.M = event_sp, 
-                                 lsoa_centr = cbind(lsoa_sp@data$centr_lng,
[…]

### 2020-05-14 src/src4_analysis.R

@@ -1,224 +1,115 @@
 
 # ANALYSIS src
-# Code to reproduce the results reported in the publication
-# source("./src/src3_recommendations.R")  # required
 
-
-# 1 SOME DESCRITPIVE STATS ---------------
+# 1  Descriptive stats ---------------
      
+  # LSOA
     length(lsoa_sp$code) # n LSOAs
     summary.stats(lsoa_sp$pop) # population
-    summary.stats(lsoa_sp$mn_dstn) # distances to the nearest event
-    summary.stats(lsoa_sp$run_count) # run counts raw
-    summary.stats(lsoa_sp$runs_pmil.week) # run rates
     summary.stats(lsoa_sp$imd_sc) # deprivation indices
+    summary.stats(lsoa_sp$mn_dstn) # distances to the nearest event
+  
+  # events
     length(event_sp$course) # n parkrun events
     summary.stats(event_sp$srvd_pop) # served population
     summary.stats(event_sp$srvd_lsoa) # served LSOA
+  
+  # green spaces    
+    summary.stats(greens_sp$area_km2) # green space areas
     
-    # pop living_within_1km 
-    round(sum(lsoa_sp$pop[lsoa_sp$mn_dstn<=1])/sum(lsoa_sp$pop),4)*100
-    # pop living_within_5km
-    round(sum(lsoa_sp$pop[lsoa_sp$mn_dstn<=5])/sum(lsoa_sp$pop),4)*100
-    # pop living_within_10km
-    round(sum(lsoa_sp$pop[lsoa_sp$mn_dstn<=10])/sum(lsoa_sp$pop),4)*100
-    # pop living_outside_25km
-    round(sum(lsoa_sp$pop[lsoa_sp$mn_dstn>25])/sum(lsoa_sp$pop),4)*100
+  # pop living_within_xkm catchment 
+    catch_pop(1)
+    catch_pop(5)
+    catch_pop(10)
+    catch_pop(20)
+    
+    # descriptive stats new distances
+    sum(lsoa_sp$mn_dstn_diff != 0) # lsoas affected
+    sum(lsoa_sp$pop[lsoa_sp$mn_dstn_diff != 0]) # pop affected
+    sum(lsoa_sp$pop[lsoa_sp$mn_dstn_diff != 0])/sum(lsoa_sp$pop) # pop %
+    summary.stats(lsoa_sp$mn_dstn_diff[lsoa_sp$mn_dstn_diff != 0])
[…]

### 2020-05-14 manuscript.Rmd
minor revisions an corrections in manuscript format tbl function kable extra table

@@
 parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professional with respect to both.
-Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2004; @wiltshire2018]
+Since its founding as a small event in London with 13 participants in 2004, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2004; @wiltshire2018]
 
@@
   This study is a cross-sectional ecological analysis of the socio-economic disparities in geographic access to parkrun events in England at the end of 2018.
-All analyses were conducted on the level of Lower Layer Super Output Areas (LSOAs), which divide England into 32,844 geographic units with, on average, have a population of approximately 1,700.
+All analyses were conducted on the level of Lower Layer Super Output Areas (LSOAs), which divide England into 32,844 geographic units which, on average, have a population of approximately 1,700.
 We assessed the relationship between access, defined as the distance (as the crow flies) to the nearest parkrun event, and socio-economic deprivation, measured using the Index of Multiple Deprivation.
-In addition, we used information on green spaces in England to conducted a simple location-allocation analysis, in order to identify 200 locations for future parkrun events, that maximise access for access for the population.
+In addition, we used information on public green spaces in England to conducted a simple location-allocation analysis, in order to identify 200 locations for future parkrun events, that maximise access for the population.
 
@@
 
-  For this study, we combined data on three types of geospatial entities: 1) LSOAs, 2) parkrun event locations, and 3) public green spaces.
+  For this study, we combined data on three types of geospatial entities: 1) LSOAs, 2) parkrun events, and 3) public green spaces.
 
-  1) The geographic locations of all 32,844 LSOAs, defined by the coordinates of its population-weighted centroid, 2017 total population estimates, and the 2015 Index of Multiple Deprivation (IMD) were retrieved from the Office for National Statistics [@ons2018; @ref; @ref].
+  1) For all 32,844 LSOAs, we retrieved geographic locations, defined by the coordinates of its population-weighted centroid, 2017 total population estimates, and the 2015 Index of Multiple Deprivation (IMD) from the Office for National Statistics [@ons2018; @ref; @ref].
 
-  2) We included all 465 public parkrun events which were in operation in England by December 12th 2018 - on this date, Sport England announced that their plan to provide funding for setting up 200 additional parkrun events across England.[@ref] The coordinates of all 465 events were manually scrapped from the parkrun UK website.[@ref]
+  2) We included all 465 public parkrun events which were in operation in England by December 12th, 2018 - on this date, Sport England announced their plan to provide funding for setting up 200 additional parkrun events across England.[@ref] The locations of the events were manually scrapped from the parkrun UK website.[@ref]
 
-  3) The location coordinates of 143,822 public green spaces were retrieved from an open dataset of Ordnance Survey.[9] parkrun events are held in various settings and terrains, and not always require a 5km loop - some events have courses that involve running back and forth.
-After evaluating existing parkrun event courses, we decided to consider all public parks, gardens, and playing fields in England with an area of 0.1 km$^2$ (e.g.
+  3) The locations of public green spaces in England were retrieved from an open dataset of Ordnance Survey.[9] Even though parkrun events do not always require a 5km loop (some courses involve running back and forth) and are held in various settings and terrains, not all green spaces are suitable to host running events.
+After evaluating existing parkrun event courses, we decided to consider all public parks, gardens, and playing fields with an area of 0.1 km$^2$ (e.g.
 316m x 316m) or more potentially suitable for hosting events (n=2,842).
@@
 
-  Access to parkrun was measured in terms of the distance (as the crow flies) to the nearest event.
-For each of the 32,844 LSOAs, we computed the geodesic distances between its population-weighted centroid and all 465 parkun events and then selected the shortest distance (in km).
+  The two variables of interest were access to parkrun events and deprivation of LSOAs.
 
-  The socio-economic deprivation of LSOAs was measured using IMD.
-It is a measure of measure of relative deprivation, which combines 37 indicators from seven domains (income, employment, education and skills, health and disability, crime, housing and services, and living environment) into a single score.
+  Access to parkrun was measured in terms of the distance as the crow flies (in km) to the nearest event.
+For each of the 32,844 LSOAs, we computed the geodesic distances between its population-weighted centroid and all 465 parkun events that were in operation on December 12th, 2018, and then selected the shortest distance.
+
+  The socio-economic deprivation of LSOAs was measured using the 2015 IMD.
+It is a measure of relative deprivation, which has been used in many similar studies.
+The IMD combines 37 indicators from seven domains (income, employment, education and skills, health and disability, crime, housing and services, and living environment) into a single score.
 The score ranges from 0 (least deprived) to 100 (most deprived).[@smith2015]
 
-  Further covariates (e.g.
-population density or demographics), which are likely to affect the availability of parkrun events, were not taken into account, because we did not aim to assess to what extent deprivation independently 'explains' access.
-Rather, we sought to evaluate whether or not people living in deprived areas actually have better or worse access, given all other circumstances.
+  Other covariates, which are likely to affect the availability of parkrun events within an area (e.g.
+population density or demographics), were not taken into account, because we did not aim to assess to what extent deprivation independently 'explained' access.
+Rather, we sought to evaluate whether or not people living in deprived areas have better or worse access, given the actual circumstances.
 
@@
   Mean, standard deviation, median, interquartile range, and range were used as descriptive statistics.
-The correlation between IMD and the distance to the nearest parkrun event on the LSOA level were investigated using the Pearson and Spearman method.
-The correlation coefficients were computed using the LSOAs' total population as weights.
-A scatter plot with a fitted linear regression model are presented to illustrate the relationship.
+We then assessed the association between the IMD and the distance to the nearest parkrun event on the LSOA level.
 Our hypothesis was that more deprived areas had worse access, i.e.
 longer distances to the nearest parkrun event than more affluent areas.
+Pearson and Spearman correlation coefficients were computed using the LSOAs' total population as weights.
+We also conducted a stratified analysis, for which we grouped LSOAs into IMD quintiles (most, more, median, less, least deprived) and assessed access to parkrun events in each strata.
 
@@
   We assessed the overall impact of setting up 200 new parkrun events at the identified optimal location on the geographic access to parkrun events in England. 
-We also conducted a distributional effectiveness analysis, in which we grouped LSOAs into IMD quintiles (most, more, median, less, least deprived) and subsequently assessed how setting up the 200 new events would affect access in each of those subgroups.
+  In order to conduct a distributional analysis, we investigated the effect of setting up the 200 new events on LSOAs across IMD quintiles.
 
@@
 
-  As of 12th December 2018, approximately 7%, 69%, and 91% of the English population lived within 1, 5, and 10km of a parkrun event.
+  As of 12th December 2018, approximately 7%, 69%, and 91% of the English population lived within 1, 5, and 10km distance of a parkrun event.
 Only 578,043 people (1% of the English population) lived more than 20km from an event.
@@
 
-| **Table 1: Descriptive statistics of LSOAs and parkrun events** |
-| ---|---|---|--- |
-| <td colspan=4>**LSOA (n =n 32,844) ** |
-| ---|---|---|--- |
-| |Mean (SD) | Median (Q1-Q3) | Range |
-| Population | 1,693.44 (405.35) | 1,612 (1,452-1,834) | 362-13,404 |
-| IMD score | 21.67 (15.59) | 17.40 (9.65-30.07) | 0.48-92.60 |
-| Distance to the nearest parkrun event location | 4.65 (4.22) | 3.39 (1.99-5.83) | 0.04-76.44 |
-|
-| <td colspan=4>**Parkrun Events (n = 465)** |
-| ---|---|---|--- |
-| Served LSOAs\* | 70.63 (42.81) | 62 (40-87) | 6-350 |
-| Served population\* | 119,612 (74,290) | 103,952 (68,837-151,488) | 7,855-628,010 |
-| ---|---|---|--- |
- | \* Number of LSOAs or population for which a given parkrun event is the nearest. |
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+load(file = "./output/desc_tbls.RDS")
+library(knitr)
+library(kableExtra)
+
+kable(rbind(desc_tbl1,desc_tbl2),
+      caption = "Table 1: Descriptive statistics of LSOAs and parkrun events",
+      format = "latex",
+      longtable = T, 
+      booktabs = T) %>%
+  group_rows(group_label = "LSOAs (n = 32,844)",
+             start_row = 1,hline_before=F,
+             end_row = 3) %>%
+   group_rows(group_label = "Parkrun events (n = 465)",
+             start_row = 4,hline_before=T,
+             end_row = 5) %>%
+  add_footnote("foot1")
+
+```
 
@@
 
-The distributional analysis of distances by IMD quantile showed that people living in the 20% most deprived LSOAs  had the best geographic access, with an average and median distance to the nearest parkrun event of 3.51 and 2.79 km.
+The distributional analysis of distances by IMD quintile showed that people living in the 20% most deprived LSOAs  had the best geographic access, with an average and median distance to the nearest parkrun event of 3.51 and 2.79 km.
 Depending on the metric, the worst access was observed for LSOAs in the middle (mean distance = 3.36km) or the less deprived group (median distance = 3.93km).
@@
 Table 2 shows the difference in geographic access before and after the introduction of the 200 new parkrun events at the identified optimal green spaces.
-Results are presented for the overall effect and stratified by IMD quantiles.
+Results are presented for the overall effect and stratified by IMD quintiles.
 The 200 new events amplified the relationship between IMD and geographic access further: the population-weighted Pearson and Spearman correlation coefficients changed from -0.15 and -0.18 before, to -0.20 and -0.23 afterwards.
-Nevertheless, the improvement in access was smallest for LSOAs in the most deprived quantile.
+Nevertheless, the improvement in access was smallest for LSOAs in the most deprived quintile.
 
@@
 # Table 2
-# quantile(lsoa_sp$imd_sc,seq(0,1,0.2))
+# quintile(lsoa_sp$imd_sc,seq(0,1,0.2))
 # cbind(mndst_tbl,mndst_new_tbl)

### 2020-05-14 src/src1_functions.R

@@ -27,6 +27,16 @@ install_n_load = function(package){
    round(res,round.dec) 
   }
   
+# format descriptive tbl 
+  format.tbl = function(desc_tbl1,cols){
+    desc_tbl1 = cbind(paste(desc_tbl1$mean.d," (",desc_tbl1$sd.d,")",sep=""),
+                      paste(desc_tbl1$median.d," (",desc_tbl1$q25.d,";",desc_tbl1$q75.d ,")",sep=""),
+                      paste(desc_tbl1$min.d,"-",desc_tbl1$max.d,sep=""))
+    rownames(desc_tbl1) = cols
+    colnames(desc_tbl1) = c("Mean (SD)","Median (Q25-Q75)","Range")
+    return(desc_tbl1)
+  }
+  
 # n % living within x km
   catch_pop = function(x,dist = lsoa_sp$mn_dstn ){
     paste(round(sum(lsoa_sp$pop[dist<=x])/sum(lsoa_sp$pop),4)*100,"/ ", formatC(sum(lsoa_sp$pop[dist<=x]),format="f",big.mark = ","))

### 2020-05-14 src/src4_analysis.R

@@ -4,16 +4,23 @@
 # 1  Descriptive stats ---------------
      
   # LSOA
-    length(lsoa_sp$code) # n LSOAs
-    summary.stats(lsoa_sp$pop) # population
-    summary.stats(lsoa_sp$imd_sc) # deprivation indices
-    summary.stats(lsoa_sp$mn_dstn) # distances to the nearest event
-  
+    length(lsoa_sp$code) # n LSOAs  
+    desc_tbl1 =rbind(
+      summary.stats(lsoa_sp$pop), # population
+      summary.stats(lsoa_sp$imd_sc), # deprivation indices
+      summary.stats(lsoa_sp$mn_dstn)) # distances to the nearest event
+    desc_tbl1 = format.tbl(desc_tbl1 = desc_tbl1,cols=c("Population","IMD","Distance to the nearest event"))
+    # desc_tbl1
+
   # events
     length(event_sp$course) # n parkrun events
-    summary.stats(event_sp$srvd_pop) # served population
-    summary.stats(event_sp$srvd_lsoa) # served LSOA
-  
+    desc_tbl2 =rbind(
+      summary.stats(event_sp$srvd_pop), # served population
+      summary.stats(event_sp$srvd_lsoa)) # served LSOA
+    desc_tbl2 = format.tbl(desc_tbl2,cols=c("Catchment area population","Catchment area LSOAs"))
+    # desc_tbl2
+    save(list= c("desc_tbl1","desc_tbl2"),file = "./output/desc_tbls.RDS")
+    
   # green spaces    
     summary.stats(greens_sp$area_km2) # green space areas
     

### 2020-05-14 manuscript.Rmd
changes to introduction

--- 2020-05-14 manuscript.Rmd
@@
 
-Insufficient physical activity is a leading cause of disease and disability worldwide.[@who2009] For the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic background are disproportionately affected by physical inactivity-related diseases.
+Insufficient physical activity is a leading cause of disease and disability worldwide.[@who2009] In the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic background are more likely to be inactive [@farrell2014socioeconomic] and are disproportionately affected by physical inactivity-related diseases [@tanaka2018socioeconomic].
 Increasing the physical activity levels of the population is therefore high on the public health agenda: it not only has the potential to improve quality of life, reduce mortality rates and alleviate the strain on health and social care services, but also reduce the gap in health inequalities.
@@
 parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professional with respect to both.
-Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2004; @wiltshire2018]
+Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass participation sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation, particularly in individuals who were previously inactive or did not meet public health guidelines.[@stevinson2004; @wiltshire2018]
 
-Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely driven by *demand*, not *need*.
-It might therefore be the case that parkrun events are primarily located in areas that are relatively well-off anyway, while people living in more deprived communities may not have the same opportunities to participate.
+Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely grassroots, driven by *demand* rather than *need*.
+It might therefore be the case that parkrun events are primarily located in areas that are relatively affluent, while people living in more deprived communities may not have the same opportunities to participate.
 In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@sportengland2018]
@@
 The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
-Secondly, to identify 200 optimal locations for future events, in order to improve geographic access for deprived communities.
+Secondly, to identify 200 optimal locations for future events in order to improve geographic access, in particular for deprived communities.
 

### 2020-05-14 manuscript.Rmd

--- 2020-05-14 manuscript.Rmd
@@
 parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professional with respect to both.
-Since its founding as a small event in London with 13 participants in 2004, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2004; @wiltshire2018]
+Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass participation sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation, particularly in individuals who were previously inactive or did not meet public health guidelines.[@stevinson2004; @wiltshire2018]
 
-Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely driven by *demand*, not *need*.
-It might therefore be the case that parkrun events are primarily located in areas that are relatively well-off anyway, while people living in more deprived communities may not have the same opportunities to participate.
+Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely grassroots, driven by *demand* rather than *need*.
+It might therefore be the case that parkrun events are primarily located in areas that are relatively affluent, while people living in more deprived communities may not have the same opportunities to participate.
 In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@sportengland2018]

### 2020-05-15 manuscript.Rmd

--- 2020-05-14 manuscript.Rmd
@@
 
-Insufficient physical activity is a leading cause of disease and disability worldwide.[@who2009] In the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic background are more likely to be inactive [@farrell2014socioeconomic] and are disproportionately affected by physical inactivity-related diseases [@tanaka2018socioeconomic].
+Insufficient physical activity is a leading cause of disease and disability worldwide.[@who2009] For the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic background are disproportionately affected by physical inactivity-related diseases.
 Increasing the physical activity levels of the population is therefore high on the public health agenda: it not only has the potential to improve quality of life, reduce mortality rates and alleviate the strain on health and social care services, but also reduce the gap in health inequalities.
@@
 
-parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professional with respect to both.
-Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass participation sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation, particularly in individuals who were previously inactive or did not meet public health guidelines.[@stevinson2004; @wiltshire2018]
+parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professionals with respect to both.
+Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2004; @wiltshire2018]
 
-Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely grassroots, driven by *demand* rather than *need*.
-It might therefore be the case that parkrun events are primarily located in areas that are relatively affluent, while people living in more deprived communities may not have the same opportunities to participate.
+Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely driven by *demand*, not *need*.
+It might therefore be the case that parkrun events are primarily located in areas that are relatively well-off anyway, while people living in more deprived communities may not have the same opportunities to participate.
 In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@sportengland2018]
@@
 The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
-Secondly, to identify 200 optimal locations for future events in order to improve geographic access, in particular for deprived communities.
+Secondly, to identify 200 optimal locations for future events, in order to improve geographic access for deprived communities.
 
@@
   This study is a cross-sectional ecological analysis of the socio-economic disparities in geographic access to parkrun events in England at the end of 2018.
-All analyses were conducted on the level of Lower Layer Super Output Areas (LSOAs), which divide England into 32,844 geographic units with, on average, have a population of approximately 1,700.
-We assessed the relationship between access, defined as the distance (as the crow flies) to the nearest parkrun event, and socio-economic deprivation, measured using the Index of Multiple Deprivation.
-In addition, we used information on green spaces in England to conducted a simple location-allocation analysis, in order to identify 200 locations for future parkrun events, that maximise access for access for the population.
+All analyses were conducted on the level of Lower layer Super Output Areas (LSOAs), which divide England into 32,844 geographic units with, on average, a population of approximately 1,700.
+We assessed the relationship between access, defined as the distance (as the crow flies) to the nearest parkrun event, and socio-economic deprivation, measured using the Index of Multiple Deprivation (IMD).
+In addition, we used information on green spaces in England to conduct a simple location-allocation analysis in order to identify 200 locations for future parkrun events that maximise access for the English population.
 
@@
 
-  1) The geographic locations of all 32,844 LSOAs, defined by the coordinates of its population-weighted centroid, 2017 total population estimates, and the 2015 Index of Multiple Deprivation (IMD) were retrieved from the Office for National Statistics [@ons2018; @ref; @ref].
+  1) The geographic locations of all 32,844 LSOAs, defined by the coordinates (longitude & latitude) of its population-weighted centroid, 2017 total population estimates, and the 2015 Index of Multiple Deprivation (IMD) were retrieved from the Office for National Statistics [@ons2018; @ref; @ref].
 
-  2) We included all 465 public parkrun events which were in operation in England by December 12th 2018 - on this date, Sport England announced that their plan to provide funding for setting up 200 additional parkrun events across England.[@ref] The coordinates of all 465 events were manually scrapped from the parkrun UK website.[@ref]
+  2) We included all 465 public parkrun events which were in operation in England by December 12th 2018, the date on which Sport England announced their plan to provide funding to set up 200 additional parkrun events across England.[@ref] The coordinates of all 465 events were obtained from the parkrun UK website.[@ref]
 
-  3) The location coordinates of 143,822 public green spaces were retrieved from an open dataset of Ordnance Survey.[9] parkrun events are held in various settings and terrains, and not always require a 5km loop - some events have courses that involve running back and forth.
-After evaluating existing parkrun event courses, we decided to consider all public parks, gardens, and playing fields in England with an area of 0.1 km$^2$ (e.g.
-316m x 316m) or more potentially suitable for hosting events (n=2,842).
+  3) The coordinates of 143,822 public green spaces were retrieved from an open Ordnance Survey dataset.[9] parkrun events are held in various settings and terrains, and do not always require a single 5km loop - some events have courses that involve running back and forth.
+After evaluating existing parkrun event courses, we decided to consider all public parks, gardens, and playing fields in England with an area of 0.1 km$^2$ or more potentially suitable for hosting events (n=2,842).
 
@@
   Access to parkrun was measured in terms of the distance (as the crow flies) to the nearest event.
-For each of the 32,844 LSOAs, we computed the geodesic distances between its population-weighted centroid and all 465 parkun events and then selected the shortest distance (in km).
+For each of the 32,844 LSOAs, we computed the geodesic distances (distance as the crow flies) between its population-weighted centroid and all 465 parkrun events and then selected the shortest distance (in km).
+This distance is the LSOA's distance to the nearest parkrun event - our measure of geographic access.
 
   The socio-economic deprivation of LSOAs was measured using IMD.
-It is a measure of measure of relative deprivation, which combines 37 indicators from seven domains (income, employment, education and skills, health and disability, crime, housing and services, and living environment) into a single score.
+It is a measure of relative deprivation, which combines 37 indicators from seven domains (income, employment, education and skills, health and disability, crime, housing and services, and living environment) into a single score.
 The score ranges from 0 (least deprived) to 100 (most deprived).[@smith2015]
@@
 population density or demographics), which are likely to affect the availability of parkrun events, were not taken into account, because we did not aim to assess to what extent deprivation independently 'explains' access.
-Rather, we sought to evaluate whether or not people living in deprived areas actually have better or worse access, given all other circumstances.
+Rather, we sought to evaluate whether or not people living in deprived areas have better or worse access, given all other circumstances.
 
@@
 The correlation coefficients were computed using the LSOAs' total population as weights.
-A scatter plot with a fitted linear regression model are presented to illustrate the relationship.
+A scatter plot with a fitted linear regression model is presented to illustrate the relationship.
 Our hypothesis was that more deprived areas had worse access, i.e.
@@
 There are 2,842 public green spaces in England in which new events could be set up.
-Which 200 locations should be selected, in order to maximise access for the greatest number of people?
+Which 200 locations should be selected, in order to maximise access for the English population?
 
@@
 Firstly, for each green space, we evaluate how setting up a parkrun event would affect the sum of distances, given the locations of all existing events (i.e.
-for how many LSOAs would this green space then be the nearest parkrun event, and by how much would it decrease the respective distances).
+for how many LSOAs would this green space then be the nearest parkrun event, and by how much an event in this location would decrease the respective distances).
 Secondly, the green space with the greatest effect is selected and added to the set of existing parkrun events.
@@
 
-  In order to identify the optimal 200 new locations for setting up new parkun events, the selection procedure is repeated 200 times.
+  In order to identify the optimal 200 new locations for setting up new parkrun events, the selection procedure is repeated 200 times.
 At each step, the single best candidate green space location is selected, added to the set of established parkrun events $E$, and removed from the set of available green spaces $C$.
-This means, the optimal green space selected at step $k$ is is taken into account when selecting the $k$th+1 location.
+This means, the optimal green space selected at step $k$ is taken into account when selecting the $k$th+1 location.
 
@@
 
-There was a negative relationship between IMD and the distance to the nearest parkrun event: the (population weighted) Pearson and Spearman correlation coefficients were -0.15 and -0.18, indicating a small effect.
-This means, more deprived LSOAs tended to have shorter distances to the nearest parkun event, i.e.
-better access.
+There was a negative relationship between IMD and the distance to the nearest parkrun event: the (population weighted) Pearson and Spearman correlation coefficients were -0.15 and -0.18, indicating a small negative correlation between deprivation and access.
+This means that on 12th December 2018, those living in more deprived LSOAs tended to have shorter geographic distances to their nearest parkrun event, i.e.
+better access, than those living in less deprived areas.
 
-The distributional analysis of distances by IMD quantile showed that people living in the 20% most deprived LSOAs  had the best geographic access, with an average and median distance to the nearest parkrun event of 3.51 and 2.79 km.
+The distributional analysis of distances by IMD quintile showed that people living in the 20% most deprived LSOAs had the best geographic access, with an average and median distance to the nearest parkrun event of 3.51 and 2.79 km.
 Depending on the metric, the worst access was observed for LSOAs in the middle (mean distance = 3.36km) or the less deprived group (median distance = 3.93km).
@@
 
-Figure 1 shows the parkrun events (black circles) that existed on 12th December 2018 alongside recommendations for 200 additional event locations (red triangles), which maximise overall access to parkrun for the greatest number of people.
+Figure 1 shows the parkrun events (black circles) that existed on 12th December 2018 alongside recommendations for 200 additional event locations (red triangles), which maximise overall access to parkrun for the English population.
 The numbers correspond to the rank, where 1 is the location which would improve access the most.
-The names and exact locations of the selected 200 green spaces are provided in the appendix.
+The names and geographic coordinates of the selected 200 green spaces are provided in the appendix.
 
@@
 **Figure 1: Map of England showing current parkrun events (blue circles) and recommended new event locations (red triangles) ranked in descending order of estimated effect on overall population-weighted access.
-Information on all 200 identified optimal green space locations are provided in the appendix.**
+Information on all 200 identified optimal green space locations is provided in the appendix.**
 
-We estimated that setting up new parkrun events in those 200 green spaces would improve access for around 16.5 million people (30% of th population) from 9,854 LSOAs.
-For those, the distance to the nearest event would, on average, be reduced by 4.09 km (SD = 3.97).
+We estimate that setting up new parkrun events in those 200 green spaces would improve access for around 16.5 million people (30% of the population) from 9,854 LSOAs.
+For these areas, the distance to the nearest event would, on average, be reduced by 4.09 km (SD = 3.97).
 The percentage of people who live within 5km of a parkrun would increase from 69% to 81.9%.
@@
 Table 2 shows the difference in geographic access before and after the introduction of the 200 new parkrun events at the identified optimal green spaces.
-Results are presented for the overall effect and stratified by IMD quantiles.
+Results are presented for the overall effect and stratified by IMD quintiles.
 The 200 new events amplified the relationship between IMD and geographic access further: the population-weighted Pearson and Spearman correlation coefficients changed from -0.15 and -0.18 before, to -0.20 and -0.23 afterwards.
@@
 ![](./output/hist.png)
-**Figure 2: caption**
+**Figure 2: Density plot of access to parkrun events before and after the creation of 200 new events**
 
@@
 
-<!-- Main findings -->
-- Access to parkrun is good
-- Contrary to our expectation, we did not find that access is better for people living in less deprived areas. On the contrary, people in the most deprived areas had the best geographic access.
-- setting up 200 new events in optimal green spaces would further improve access for a large part of the population. 
-- Setting up new events in the way we proposed would again benefit
+Overall, geographic access to parkrun in England is good.
+On 12th December 2018 more than two thirds of the population lived within 5km (the parkrun distance) of a parkrun event.
+Contrary to our expectation, we did not find that access was better for people living in more affluent areas.
+In fact, those living in the most deprived areas had the best geographic access to parkrun.
 
-<!-- further considerations -->
-- an explanation for the negative association between IMD and distance is urban/rural: deprived areas tend to cluster in cities and this is also where most parkruns take place
+Our analysis has shown that setting up 200 new events in the optimal green spaces in England would increase the percentage of English residents who live within 5k of a parkrun to 82%.
+This would increase the inequity in access to parkrun events, improving the geographic access of the most deprived more than the access of those living in more affluent areas.
+It is rare in public health for inequity to exist in this direction.
 
-<!-- link to relevant studies -->
+While our analysis has not been concerned with confounding factors, we are interested only in geographic access based on multiple deprivation, an explanation for the negative relationship between IMD and access is population density.
+Since deprived areas cluster in cities, geographical access as the crow flies tends to be better for these areas in part because population density is itself a predictor of the existence of parkrun events.
 
-<!-- strengths & limitations -->
-- first quantitative analysis
-- geographic access is not travel distance or travel time
-- greenspace: not all are suitable, and some suitable places (promenades, beaces, national parks) are not included.
-However, green spaces are distrubuted over England and it seems unlikely that any important area was missed completely.
+There are several strengths of this study, it is the first study of geographic access to parkrun in England - therefore the approach is novel and the data untapped.
+Secondly the analysis makes use of large and rich datasets, with over 30,000 LSOA and over 400 existing parkrun events it is unlikely that individual outliers are affecting the results.
+However, there are also limitations.
+Most importantly geographic access is not travel distance, or travel time.
+In some cases, for example where natural barriers like lakes or rivers block routes, the actual distance travelled may be far in excess of the distance as the crow flies.
+Also, in rural areas where car use is more prevalent travel time may be shorter, despite travel distances being longer.
+Further investigation could determine how perceptions of access change with distance in different settings (e.g.
+rural vs urban).
 
-<!-- main discussion points -->
-1 ) - proximity only provides the opportunity to participate in parkrun events
-  - in a previous study we found that after controlling for distance and other factors, people from deprived areas are less likely to participate in parkrun events.
-  - this means, giving access is neccessary but not sufficient to increase participation from deprivied communities.
-  - parkrun should complement their efforts with...
+The choice of 0.1sq-km for greenspaces may also surprise some readers.
+However, several parkruns exist in very small areas (Mile End park in London, and Cudmore Grove park in Essex are 0.14sq-km and 0.12sq-km respectively).
+We therefore chose to maximize sensitivity, at the likely expense of specificity, and therefore it is possible that some of the recommendations are in parks that are not suitable to host a parkrun event.
+Conversely, many blue spaces such as beaches and promenades are not included in the database and therefore there may exist some locations which we have missed, particularly in coastal areas.
 
-2 ) - future research should build on this work and develop a model to assess the (cost-)effectiveness of setting up new events not only in terms of improved potential access, but actual participation.
-This requires estimating causal and marginal effects.
-More, longitudinal data and more sophisticated modelling are required.
+The main finding, that geographical access to parkrun events is better in more deprived communities, is surprising given that we know that participation in parkrun has a strong socioeconomic gradient with higher participation rates in more affluent areas [@smithschneider].
+This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been enough to engage deprived communities and reduce inequity.
 
-<!-- end on positive note -->
+Investigating other barriers to participation in parkrun is likely to improve our understanding of the reasons why physical activity levels are lower in more deprived areas in general, and give parkrun the information it needs to develop strategies, other than improving access, to engage those who live in deprived communities.
+Future research should build on this work and develop a model to assess the (cost-)effectiveness of setting up new events, and other strategies, not only in terms of improved potential access, but actual participation.
+This requires estimating the causal and marginal effects of different interventions on participation, and therefore physical activity levels, using longitudinal data and sophisticated modelling techniques.
 
-- parkrun is hugely succesfull and has the potential to improve levels of physcial activity of the pop?
+parkrun has been hugely successful in encouraging physical activity worldwide, it will achieve its vision of a 'healthier, happier planet' [@parkrun2017], but the extent of its impact will be dependent on how well it is able to understand the drivers of participation and develop strategies to overcome the hurdles it identifies.
 
@@
 
-## References
-
-<div id="refs"></div>
-
-## Appendix
+Geographic access to parkrun is good, but can be improved considerably with the creation of 200 new events.
+Surprisingly, we find that geographic access is better for those living in more deprived communities.
+Given that we know participation rates are much lower in deprived areas, it seems likely that improving access alone is unlikely to significantly reduce inequity in participation.
+A deeper understanding is needed of the barriers to participation in deprived communities if strategies are going to be created to improve engagement.

### 2020-05-15 manuscript.Rmd
* new table formatting * k.step change plot * distr. analysis * code reformatted

--- 2020-05-14 manuscript.Rmd
@@
 
-  The function yields the candidate green space $c$, from the set of all 2,842 green spaces $C$, which minimises the sum of the products of the minimum distances and the total population of all LSOAs, whereby $p_i$ denotes the population of LSOA $i$, and $d_i(E \cap c)$ denotes the distance to the nearest parkrun event, which can either be an existing event from the set $E = \{e_{1},e_{2},...,e_{465}\}$, or the candidate green space $c$, whichever is nearest.
+  The function yields the candidate green space $c$, from the set of all 2,842 green spaces $C$, which minimises the sum of the population-weighted distances between LSOAs and their nearest parkrun event.
+The total population of LSOA $i$ is denoted $p_i$, and $d_i(E \cap c)$ denotes LSOA $i$'s distance to the nearest parkrun event, which can either be an existing event from the set of 465 parkrun events, denoted $E$, or the candidate green space $c$, whichever is nearest.
 
-  In order to identify the optimal 200 new locations for setting up new parkun events, the selection procedure is repeated 200 times.
+  In order to identify the optimal new locations for setting up 200 new parkun events consecutively, the selection procedure is repeated 200 times.
 At each step, the single best candidate green space location is selected, added to the set of established parkrun events $E$, and removed from the set of available green spaces $C$.
-This means, the optimal green space selected at step $k$ is is taken into account when selecting the $k$th+1 location.
+This means, the effect of the green space selected at step $k$ is taken into account when selecting the $k$th+1 location.
 
-#### Evaluating the location allocation analysis
-
-  We assessed the overall impact of setting up 200 new parkrun events at the identified optimal location on the geographic access to parkrun events in England. 
-  In order to conduct a distributional analysis, we investigated the effect of setting up the 200 new events on LSOAs across IMD quintiles.
+  We assessed the overall impact of setting up 200 new parkrun events on the geographic access to parkrun events in England.
+We also investigated the effects on LSOAs across IMD quintiles in a  distributional analysis.
 
@@
 
-kable(rbind(desc_tbl1,desc_tbl2),
-      caption = "Table 1: Descriptive statistics of LSOAs and parkrun events",
+kable(rbind(desc_tbl1,desc_tbl2),format.args = list(big.mark=","),
+      caption = "Descriptive statistics of LSOAs and parkrun events",
       format = "latex",
-      longtable = T, 
+      # longtable = T, 
       booktabs = T) %>%
@@
              end_row = 5) %>%
-  add_footnote("foot1")
-
+  add_footnote("Number of LSOAs/total population for which a given parkrun event is the nearest.",notation = "symbol")
 ```
@@
 For those, the distance to the nearest event would, on average, be reduced by 4.09 km (SD = 3.97).
+Overall, it would reduce the average and median distance to the nearest parkrun event from 4.65 and 3.39 km to 3.43 and 2.59 km.
 The percentage of people who live within 5km of a parkrun would increase from 69% to 81.9%.
+Figure 2 shows the distribution of distances between LSOAs and their nearest parkrun events before and after the introduction of the 200 new optimal event locations at the identified optimal green spaces.
 
-Table 2 shows the difference in geographic access before and after the introduction of the 200 new parkrun events at the identified optimal green spaces.
-Results are presented for the overall effect and stratified by IMD quintiles.
-The 200 new events amplified the relationship between IMD and geographic access further: the population-weighted Pearson and Spearman correlation coefficients changed from -0.15 and -0.18 before, to -0.20 and -0.23 afterwards.
-Nevertheless, the improvement in access was smallest for LSOAs in the most deprived quintile.
+![Figure 2](./output/hist.png)
 
-Figure 2 shows the distribution of the nearest distances between LSOAs and parkrun events before and after the introduction of the 200 new optimal event locations.
+**Figure 2: Density plot showing the distance, as the crow flies, to parkrun events before and after the creation of 200 new events in identified green-spaces. **
 
-![](./output/hist.png)
-**Figure 2: caption**
+Table 2 shows further results for the effects on LSOAs across IMD quintiles.
+Overall, setting up 200 new events in the recommended green spaces would amplify the negative association between IMD and geographic access: the population-weighted Pearson and Spearman correlation coefficients changed from -0.15 and -0.18 before, to -0.20 and -0.23 afterwards, indicating that improvements in access to parkrun events were greater for more deprived LSOAs as compared to less deprived LSOAs.
+Nevertheless, the distributional analysis showed that the improvement in access was smallest for LSOAs in the most deprived quintile.
 
-```{r}
-# Table 2
-# quintile(lsoa_sp$imd_sc,seq(0,1,0.2))
-# cbind(mndst_tbl,mndst_new_tbl)
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+load(file = "./output/distr.res.RDS")
+library(knitr)
+library(kableExtra)
+
+kable(distr.res,
+      caption = "Distributional analysis",
+      format = "latex",
+      # longtable = T, 
+      booktabs = T) %>%
+  kableExtra::add_header_above(header = c(" "=1,"Current situation (December 12th, 2018)"=3,"After 200 new parkrun events are set up"=3)) %>%
+  kableExtra::kable_styling(latex_options = "scale_down") %>%
+  kableExtra::row_spec(row = 5,hline_after = T)
 ```
@@
 
-<!-- Main findings -->
+Overall, geographic access to parkrun in England is good.
+On 12th December 2018 more than two thirds of the population lived within 5km (the parkrun distance) of a parkrun event.
+Contrary to our expectation, we did not find that access was better for people living in more affluent areas.
+In fact, those living in the most deprived areas had the best geographic access to parkrun.
+
+Our analysis has shown that setting up 200 new events in the optimal green spaces in England would increase the percentage of English residents who live within 5k of a parkrun to 82%.
+This would increase the inequity in access to parkrun events, improving the geographic access of the most deprived more than the access of those living in more affluent areas.
+It is rare in public health for inequity to exist in this direction.
+
+While our analysis has not been concerned with confounding factors, we are interested only in geographic access based on multiple deprivation, an explanation for the negative relationship between IMD and access is population density.
+Since deprived areas cluster in cities, geographical access as the crow flies tends to be better for these areas in part because population density is itself a predictor of the existence of parkrun events.
+
+There are several strengths of this study, it is the first study of geographic access to parkrun in England - therefore the approach is novel and the data untapped.
+Secondly the analysis makes use of large and rich datasets, with over 30,000 LSOA and over 400 existing parkrun events it is unlikely that individual outliers are affecting the results.
+
+However, there are also limitations.
+Most importantly geographic access is not travel distance, or travel time.
+In some cases, for example where natural barriers like lakes or rivers block routes, the actual distance travelled may be far in excess of the distance as the crow flies.
+Also, in rural areas where car use is more prevalent travel time my be shorter, despite travel distances being longer.
+Further investigation could determine how perceptions of access change with distance in different settings (e.g.
+rural vs urban).
+
+The choice of 0.1sq-km for greenspaces may also surprise some readers.
+However, several parkruns exist in very small areas (Mile End park in London, and Cudmore Grove park in Essex are 0.14sq-km and 0.12sq-km respectively).
+We therefore chose to maximize sensitivity, at the likely expense of specificity, and therefore it is possible that some of the recommendations are in parks that are not suitable to host a parkrun event.
+Conversely, many blue spaces such as beaches and promenades are not included in the database and therefore there may exist some locations which we have missed, particularly in coastal areas.
+
+The main finding, that geographical access to parkrun events is better in more deprived communities, is surprising given that we know that participation in parkrun has a strong socioeconomic gradient with higher participation rates in more affluent areas (ref).
+This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been enough to engage deprived communities and reduce inequity.
+
+Investigating other barriers to participation in parkrun is likely to improve our understanding of the reasons why physical activity levels are lower in more deprived areas in general, and give parkrun the information it needs to develop strategies, other than improving access, to engage those who live in deprived communities.
+Future research should build on this work and develop a model to assess the (cost-)effectiveness of setting up new events, and other strategies, not only in terms of improved potential access, but actual participation.
+This requires estimating the causal and marginal effects of different interventions on participation, and therefore physical activity levels, using longitudinal data and sophisticated modelling techniques.
+
+parkrun has been hugely successful in encouraging physical activity worldwide, it will achieve its vision of a ‘healthier, happier planet’, but the extent of its impact will be dependent on how well it is able to understand the drivers of participation and develop strategies to over-come the hurdles it identifies.
+
+<!-- Main findings 
 - Access to parkrun is good
@@
 
-<!-- further considerations -->
+further considerations 
 - an explanation for the negative association between IMD and distance is urban/rural: deprived areas tend to cluster in cities and this is also where most parkruns take place
 
-<!-- link to relevant studies -->
+link to relevant studies
 
-<!-- strengths & limitations -->
+trengths & limitations
 - first quantitative analysis
@@
 
-<!-- main discussion points -->
+main discussion points 
 1 ) - proximity only provides the opportunity to participate in parkrun events
@@
 
-<!-- end on positive note -->
+end on positive note
 
 - parkrun is hugely succesfull and has the potential to improve levels of physcial activity of the pop?
+-->
 

### 2020-05-15 src/src1_functions.R

@@ -1,6 +1,6 @@
 
 
-# FUNCTIONS
+# SRC 1: FUNCTIONS
   
 # convenient function to install (if neccessary) and load required packages
 install_n_load = function(package){
@@ -24,13 +24,15 @@ install_n_load = function(package){
                      min.d = min(x),
                      max.d = max(x)
     )
-   round(res,round.dec) 
+    res  = c(unlist(lapply(res,function(x){formatC(x,format="f",big.mark = ",",digits = round.dec)})))
+    res = data.frame(t(res))
+    return(res)
   }
   
 # format descriptive tbl 
   format.tbl = function(desc_tbl1,cols){
     desc_tbl1 = cbind(paste(desc_tbl1$mean.d," (",desc_tbl1$sd.d,")",sep=""),
-                      paste(desc_tbl1$median.d," (",desc_tbl1$q25.d,";",desc_tbl1$q75.d ,")",sep=""),
+                      paste(desc_tbl1$median.d," (",desc_tbl1$q25.d,"; ",desc_tbl1$q75.d ,")",sep=""),
                       paste(desc_tbl1$min.d,"-",desc_tbl1$max.d,sep=""))
     rownames(desc_tbl1) = cols
     colnames(desc_tbl1) = c("Mean (SD)","Median (Q25-Q75)","Range")
@@ -131,7 +133,9 @@ install_n_load = function(package){
                 ylim=c( lat1-space.bottom,lat2+space.top)) +
       # remove labels
       xlab("") +
-      ylab("") 
+      ylab("")  +
+      theme(axis.text =  element_blank(),
+            axis.ticks = element_blank())
     
     return(static_map)
   }
@@ -145,6 +149,7 @@ install_n_load = function(package){
       name.park[is.na(name.park)] = "unnamed"
       legend_m1 = cbind("Pos"=LocAloAlg_res$pos,
                         "Park"=name.park,
+                        "Wtd. average distance"=round(LocAloAlg_res$objective/1000,2),
                         "Change (%)"=LocAloAlg_res$change,
                         "Area (km2)"=formatC(LocAloAlg_res$area_km2,digits = 2, format = "f"),
                         "Longitude"=round(LocAloAlg_res$lon,digits),

### 2020-05-15 src/src2_data_proc.R

@@ -1,8 +1,7 @@
 
-# LOAD, MERGE, AND COMPUTE THE REQUIRED DATA
+# SRC 2: Load and process data
 
-# LOAD FUNCTIONS + PACKAGES
-  # Install and load all required packages
+# Load pkgs -------
     install_n_load(c("raster",
                      "sp",
                      "sf",
@@ -22,7 +21,7 @@
                      "wCorr")
                    )
   
-# LOAD RAW DATA
+# Load raw data -------
     
     # LSOA population-weighted centroids
       # Contains public sector information licensed under the Open Government Licence v3.
@@ -60,7 +59,7 @@
       event_sp = read.csv("./input/parkrun_data/event_info_20181212.csv", stringsAsFactors = F)
       
     
-# PROCESS DATA
+# Process and combine data --------
   
   # LSOA DATA 
       # location

### 2020-05-15 src/src3_localoalg.R

@@ -93,7 +93,7 @@
   t2-t1
   # prepare results 
     top_consecutive_parks = greens_sp@data[res_df$index,]
-    top_consecutive_parks$objective = res_df$objective[-1]/1e+9 # in millions
+    top_consecutive_parks$objective = res_df$objective[-1]/sum(lsoa.pop)
     top_consecutive_parks$change = res_df$change[-1]
     top_consecutive_parks$lon = candidates.coord[res_df$index,1]
     top_consecutive_parks$lat = candidates.coord[res_df$index,2]
@@ -101,17 +101,39 @@
   
     write.csv(top_consecutive_parks,"./output/optimal_locations.csv",row.names = F)
 
+    
+# Track change in pop. weighted average distance over steps 1:200
+    obj.changes = res_df$objective/sum(lsoa.pop)
+    obj.changes = round(obj.changes/1000,8)
+    k.step = 0:top_n
+    k.change = c()
+    for(i in 2:length(obj.changes)){
+      k.change = c(k.change,obj.changes[i]-obj.changes[i-1])
+    }
+    
+    k.step.plot = plot_grid(
+                    ggplot()+
+                    geom_point(aes(x=k.step,y=obj.changes)) +
+                    geom_line(aes(x=k.step,y=obj.changes)) +
+                    theme_minimal()+
+                    ggtitle("Pop. weighted average distance to the nearest parkrun event"),
+                  ggplot()+
+                    geom_point(aes(x=k.step[-1],y=k.change)) +
+                    geom_line(aes(x=k.step[-1],y=k.change)) +
+                    theme_minimal()+
+                    ggtitle("Difference from previous step"),
+                  nrow = 1
+                    )
+    
+    ggsave(k.step.plot,filename = "./output/kstep_change.png",width = 10,height = 5)
 
     
-    
-    
-    
-# / -------
 # 4 CREATE STATIC MAPS FOR PUBLICATION -------
-# the functions below use google maps API to retrieve the baseline tile
-# to use this function, you need a api key
-# For more details, see: https://developers.google.com/maps/documentation/maps-static/intro
+  # the functions below use google maps API to retrieve the baseline tile
+  # to use this function, you need a api key
+  # For more details, see: https://developers.google.com/maps/documentation/maps-static/intro
     if(file.exists(".key.txt")){
+      # I store my api key locally in a '.key.txt' file
[…]

### 2020-05-15 src/src4_analysis.R

@@ -1,44 +1,74 @@
 
-# ANALYSIS src
+# SRC 4: ANALYSIS
 
-# 1  Descriptive stats ---------------
+# Descriptive stats 1 ---------------
      
   # LSOA
     length(lsoa_sp$code) # n LSOAs  
+    mndt_overall_tbl = summary.stats(lsoa_sp$mn_dstn,round.dec = 2)
     desc_tbl1 =rbind(
-      summary.stats(lsoa_sp$pop), # population
+      summary.stats(lsoa_sp$pop,round.dec = 0), # population
       summary.stats(lsoa_sp$imd_sc), # deprivation indices
-      summary.stats(lsoa_sp$mn_dstn)) # distances to the nearest event
+      mndt_overall_tbl) # distances to the nearest event
     desc_tbl1 = format.tbl(desc_tbl1 = desc_tbl1,cols=c("Population","IMD","Distance to the nearest event"))
     # desc_tbl1
 
   # events
     length(event_sp$course) # n parkrun events
     desc_tbl2 =rbind(
-      summary.stats(event_sp$srvd_pop), # served population
-      summary.stats(event_sp$srvd_lsoa)) # served LSOA
-    desc_tbl2 = format.tbl(desc_tbl2,cols=c("Catchment area population","Catchment area LSOAs"))
+      summary.stats(event_sp$srvd_pop,round.dec = 0), # served population
+      summary.stats(event_sp$srvd_lsoa,round.dec = 0)) # served LSOA
+    desc_tbl2 = format.tbl(desc_tbl2,cols=c("Catchment area* population","Catchment area* LSOAs"))
     # desc_tbl2
     save(list= c("desc_tbl1","desc_tbl2"),file = "./output/desc_tbls.RDS")
     
   # green spaces    
     summary.stats(greens_sp$area_km2) # green space areas
     
-  # pop living_within_xkm catchment 
+    # descriptive stats new distances
+    sum(lsoa_sp$mn_dstn_diff != 0) # lsoas affected
+    sum(lsoa_sp$pop[lsoa_sp$mn_dstn_diff != 0]) # pop affected
+    sum(lsoa_sp$pop[lsoa_sp$mn_dstn_diff != 0])/sum(lsoa_sp$pop) # pop %
+    mndst_new.tbl = summary.stats(lsoa_sp$mn_dstn_new)
[…]

### 2020-05-15 manuscript.Rmd
discussion and bib from other branch some minor revisions

@@
 
-parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professional with respect to both.
-Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass participation sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation, particularly in individuals who were previously inactive or did not meet public health guidelines.[@stevinson2004; @wiltshire2018]
+parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professionals with respect to both.
+Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2004; @wiltshire2018]
 
-Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely grassroots, driven by *demand* rather than *need*.
-It might therefore be the case that parkrun events are primarily located in areas that are relatively affluent, while people living in more deprived communities may not have the same opportunities to participate.
+Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely driven by *demand*, not *need*.
+It might therefore be the case that parkrun events are primarily located in areas that are relatively well-off anyway, while people living in more deprived communities may not have the same opportunities to participate.
 In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@sportengland2018]
 
-The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
-Secondly, to identify 200 optimal locations for future events in order to improve geographic access, in particular for deprived communities.
+The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England igeographic access to parkrun events in England is equitable across areas with different levels of deprivation.
+Secondly, to identify 200 optimal locations for future events in order geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
+Secondly, to identify 200 optimal locations for future events in order improve geographic access, in particular for deprived communities.
 
@@
 kable(distr.res,
-      caption = "Distributional analysis",
+      caption = "Distributional analysis. The table shows the access to parkrun events before and after 200 new parkrun events are set up at optimal green spaces, stratified by IMD quintiles",
       format = "latex",
@@
 In fact, those living in the most deprived areas had the best geographic access to parkrun.
+It is rare in public health for inequalities to exist in this direction.
 
-Our analysis has shown that setting up 200 new events in the optimal green spaces in England would increase the percentage of English residents who live within 5k of a parkrun to 82%.
-This would increase the inequity in access to parkrun events, improving the geographic access of the most deprived more than the access of those living in more affluent areas.
-It is rare in public health for inequity to exist in this direction.
+Our analysis has shown that setting up 200 new events in the recommended (optimal) green spaces in England would considerably improve access - for example, the percentage of English residents who live within 5km of a parkrun would increase to 82%.
+Moreover, the recommended expansion of parkrun would improve the geographic access for the most deprived communities more than the access of those living in more affluent areas.
 
-While our analysis has not been concerned with confounding factors, we are interested only in geographic access based on multiple deprivation, an explanation for the negative relationship between IMD and access is population density.
-Since deprived areas cluster in cities, geographical access as the crow flies tends to be better for these areas in part because population density is itself a predictor of the existence of parkrun events.
+Our analysis has not been concerned with confounding factors, as we were interested only in determining to what extent deprived communities have access to parkrun events - not what factors independently explain access more generally.
+It should be noted, however, that a contributing factor for the negative relationship between IMD and access is likely to be population density: deprived areas cluster in ubran areas, where also most parkun events take place.
+Rural areas, on the other hand, generally have worse geographical access as the crow flies.
 
-There are several strengths of this study, it is the first study of geographic access to parkrun in England - therefore the approach is novel and the data untapped.
+There are several strengths of this study that derserve metion.
+Firstly, it is the first study of geographic access to parkrun in England - therefore the approach is novel and the data untapped.
 Secondly the analysis makes use of large and rich datasets, with over 30,000 LSOA and over 400 existing parkrun events it is unlikely that individual outliers are affecting the results.
-
 However, there are also limitations.
-Most importantly geographic access is not travel distance, or travel time.
+Most importantly geographic access is not measured as travel distance, or travel time.
 In some cases, for example where natural barriers like lakes or rivers block routes, the actual distance travelled may be far in excess of the distance as the crow flies.
-Also, in rural areas where car use is more prevalent travel time my be shorter, despite travel distances being longer.
+Also, in rural areas where car use is more prevalent travel time may be shorter, despite travel distances being longer.
 Further investigation could determine how perceptions of access change with distance in different settings (e.g.
@@
 
-The main finding, that geographical access to parkrun events is better in more deprived communities, is surprising given that we know that participation in parkrun has a strong socioeconomic gradient with higher participation rates in more affluent areas (ref).
-This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been enough to engage deprived communities and reduce inequity.
+The main finding, that geographical access to parkrun events is better in more deprived communities, is surprising given that in a previous study we found that participation in parkrun has a strong socioeconomic gradient with considerably higher participation rates in more affluent areas.
+About a third of all participants came from LSOAs in the least deprived quintile, while only a third came from the most deprived quintile.[@smithschneider; @schneidersmith] This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been sufficient to engage deprived communities.
 
@@
 
-parkrun has been hugely successful in encouraging physical activity worldwide, it will achieve its vision of a ‘healthier, happier planet’, but the extent of its impact will be dependent on how well it is able to understand the drivers of participation and develop strategies to over-come the hurdles it identifies.
-
-<!-- Main findings 
-- Access to parkrun is good
-- Contrary to our expectation, we did not find that access is better for people living in less deprived areas. On the contrary, people in the most deprived areas had the best geographic access.
-- setting up 200 new events in optimal green spaces would further improve access for a large part of the population. 
-- Setting up new events in the way we proposed would again benefit
-
-further considerations 
-- an explanation for the negative association between IMD and distance is urban/rural: deprived areas tend to cluster in cities and this is also where most parkruns take place
-
-link to relevant studies
-
-trengths & limitations
-- first quantitative analysis
-- geographic access is not travel distance or travel time
-- greenspace: not all are suitable, and some suitable places (promenades, beaces, national parks) are not included.
-However, green spaces are distrubuted over England and it seems unlikely that any important area was missed completely.
-
-main discussion points 
-1 ) - proximity only provides the opportunity to participate in parkrun events
-  - in a previous study we found that after controlling for distance and other factors, people from deprived areas are less likely to participate in parkrun events.
-  - this means, giving access is neccessary but not sufficient to increase participation from deprivied communities.
-  - parkrun should complement their efforts with...
-
-2 ) - future research should build on this work and develop a model to assess the (cost-)effectiveness of setting up new events not only in terms of improved potential access, but actual participation.
-This requires estimating causal and marginal effects.
-More, longitudinal data and more sophisticated modelling are required.
-
-end on positive note
-
-- parkrun is hugely succesfull and has the potential to improve levels of physcial activity of the pop?
--->
+parkrun's stated mission is to create a 'healthier, happier planet' and so far they have been hugely successful in encouraging physical activity worldwide.[@parkrun2017] Our analysis has furthermore shown that parkrun events are, in principle, available to anyone, irrespective of their socio-economic background.
+However, the extent of parkrun's overall impact on health inequalities will be dependent on how well they are able to understand and overcome barriers to participation for deprived communties.
 
 ## Conclusion
+
+  In England, geographic access to parkrun events is good, and can be further improved with the creation of 200 new events.
+Contrary to our expectation, we find that geographic access is better for those living in more deprived communities.
+Given that participation rates are generally lower in deprived areas, improving access alone is unlikely to significantly reduce inequalities in participation and/or levels of physical activity.
+A deeper understanding is needed of the barriers to participation in deprived communities if strategies are going to be created to improve engagement.
 

### 2020-05-15 src/src4_analysis.R

@@ -97,8 +97,8 @@
     # 99 rows removed
     before_after_dist = ggplot(lsoa_sp@data) +
       geom_density(aes(mn_dstn_new,fill="After 200 new parkrun events are set up"),col="black",alpha=0.5) +
-      geom_density(aes(mn_dstn,fill="Current situation (December 12th, 2018)"),col="black",alpha=0.3) +
-      scale_fill_manual(values = c("Current situation (December 12th, 2018)" = "red","After 200 new parkrun events are set up"="lightgreen"),name="") +
+      geom_density(aes(mn_dstn,fill="Current access to parkrun events (December 12th, 2018)"),col="black",alpha=0.3) +
+      scale_fill_manual(values = c( "Current access to parkrun events (December 12th, 2018)" = "red","After 200 new parkrun events are set up"="lightgreen"),name="") +
       scale_x_continuous(label=label_number(suffix = " km"),limits = c(0,25),
                          name = "Distance to the nearest parkrun event") +
       theme_minimal()+

### 2020-05-15 manuscript.Rmd
minor revisions revise map legend add appendix

@@
 
-The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England igeographic access to parkrun events in England is equitable across areas with different levels of deprivation.
+The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
 Secondly, to identify 200 optimal locations for future events in order geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
-Secondly, to identify 200 optimal locations for future events in order improve geographic access, in particular for deprived communities.
+Secondly, to identify 200 optimal locations for future events in order to improve geographic access, in particular for deprived communities.
 
@@
 We assessed the relationship between access, defined as the distance (as the crow flies) to the nearest parkrun event, and socio-economic deprivation, measured using the Index of Multiple Deprivation.
-In addition, we used information on public green spaces in England to conducted a simple location-allocation analysis, in order to identify 200 locations for future parkrun events, that maximise access for the population.
+In addition, we used information on public green spaces in England to conduct a simple location-allocation analysis, in order to identify 200 locations for future parkrun events, that maximise access for the population.
 
@@
 
-Overall, geographic access to parkrun in England is good.
-On 12th December 2018 more than two thirds of the population lived within 5km (the parkrun distance) of a parkrun event.
+Overall, geographic access to parkrun in England is good. On 12th December 2018 
+The median distance to the nearest parkrun event was 3.39 km and more than two thirds of the population lived within 5km (the parkrun distance) of a parkrun event.
 Contrary to our expectation, we did not find that access was better for people living in more affluent areas.
@@
 
-Our analysis has not been concerned with confounding factors, as we were interested only in determining to what extent deprived communities have access to parkrun events - not what factors independently explain access more generally.
-It should be noted, however, that a contributing factor for the negative relationship between IMD and access is likely to be population density: deprived areas cluster in ubran areas, where also most parkun events take place.
-Rural areas, on the other hand, generally have worse geographical access as the crow flies.
+The main finding, that geographical access to parkrun events is better in more deprived communities, is surprising.
+Parkrun events are set up by volunteers, based on demand, not need.
+Studies have shown that the level of physical activity, and the availability of physical activity facilities generally declines with the level of deprivation.
+Opportunities for physical activity are often lacking in areas in most need.[e.g.
+@hillsdon2007] parkrun events, in contrast, seem to be often held in or near deprived areas, and are free to attend, giving anyone equal access, irrespective of their socio-economic background.
+Nevertheless, in a previous analysis, we found that participation in parkrun has a strong socioeconomic gradient with considerably higher participation rates in more affluent areas: about a third of all participants came from LSOAs in the least deprived quintile, while only 7% came from the most deprived quintile.[@smithschneider; @schneidersmith] This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been sufficient to engage deprived communities.
 
-There are several strengths of this study that derserve metion.
+There are several strengths of this study that deserve mention.
 Firstly, it is the first study of geographic access to parkrun in England - therefore the approach is novel and the data untapped.
 Secondly the analysis makes use of large and rich datasets, with over 30,000 LSOA and over 400 existing parkrun events it is unlikely that individual outliers are affecting the results.
+
 However, there are also limitations.
@@
 In some cases, for example where natural barriers like lakes or rivers block routes, the actual distance travelled may be far in excess of the distance as the crow flies.
-Also, in rural areas where car use is more prevalent travel time may be shorter, despite travel distances being longer.
-Further investigation could determine how perceptions of access change with distance in different settings (e.g.
-rural vs urban).
 
-The choice of 0.1sq-km for greenspaces may also surprise some readers.
-However, several parkruns exist in very small areas (Mile End park in London, and Cudmore Grove park in Essex are 0.14sq-km and 0.12sq-km respectively).
-We therefore chose to maximize sensitivity, at the likely expense of specificity, and therefore it is possible that some of the recommendations are in parks that are not suitable to host a parkrun event.
-Conversely, many blue spaces such as beaches and promenades are not included in the database and therefore there may exist some locations which we have missed, particularly in coastal areas.
+Furthermore, the list of green spaces that we considered as potential sites for future parkrun events is neither comprehensive nor without limitations.
+Not all included green spaces may be suitable to host events (e.g.
+because of the terrain or the setting), and the list also does not contain all suitable places (e.g.
+many blue spaces such as beaches and promenades are not included).
 
-The main finding, that geographical access to parkrun events is better in more deprived communities, is surprising given that in a previous study we found that participation in parkrun has a strong socioeconomic gradient with considerably higher participation rates in more affluent areas.
-About a third of all participants came from LSOAs in the least deprived quintile, while only a third came from the most deprived quintile.[@smithschneider; @schneidersmith] This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been sufficient to engage deprived communities.
+Finally.
+our analysis has been concerned only with determining to what extent deprived communities have geographic access to parkrun events.
+We did not investigate what other factors independently explain access more generally.
+It should be noted, however, that a contributing factor for the negative relationship between IMD and access is likely to be population density: deprived areas cluster in urban areas, where also most parkun events take place.
+Rural areas, on the other hand, may therefore have worse geographical access as the crow flies.
+Further studies are required to better understand wider determinants of access to parkrun and/or physical activity facilities more generally.
 
-Investigating other barriers to participation in parkrun is likely to improve our understanding of the reasons why physical activity levels are lower in more deprived areas in general, and give parkrun the information it needs to develop strategies, other than improving access, to engage those who live in deprived communities.
+  In England, parkrun has been hugely successful in engaging large numbers of people in physical activity.
+The inclusive culture of their events, and the almost universal availability of parkrun events throughout the country provide a learning opportunity to explore socio-spatial determinants influencing physical activity behaviour on a national scale.
+Our study contributes to the limited research in this area and identifies possible leads for further investigations.
+Studying barriers to participation in parkrun, other than geographic access, is likely to improve our understanding of the reasons why physical activity levels are lower in more deprived areas and may help to design more effective public health interventions to increase levels of physical activity in the population.
 Future research should build on this work and develop a model to assess the (cost-)effectiveness of setting up new events, and other strategies, not only in terms of improved potential access, but actual participation.
 This requires estimating the causal and marginal effects of different interventions on participation, and therefore physical activity levels, using longitudinal data and sophisticated modelling techniques.
-
-parkrun's stated mission is to create a 'healthier, happier planet' and so far they have been hugely successful in encouraging physical activity worldwide.[@parkrun2017] Our analysis has furthermore shown that parkrun events are, in principle, available to anyone, irrespective of their socio-economic background.
-However, the extent of parkrun's overall impact on health inequalities will be dependent on how well they are able to understand and overcome barriers to participation for deprived communties.
 
@@
 
+\newpage
+
 ## Appendix
+
+```{r echo=FALSE}
+
+  kable(read.csv("./output/map_legend.csv"),
+        caption = "List of all 200 recommended green spaces for setting up parkrun events to maximise geographic access.",
+        format = "latex",align = "lcccccc",
+        col.names = linebreak(c("Park name",
+                                "Total wtd. average distance",
+                                "% Change from previous step",
+                                "Park area (km2)",
+                                "Longitude",
+                                "Lattitude")),
+        longtable = T, 
+        booktabs = T) %>%
+  kable_styling(font_size = 7,full_width = F) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(2:4, width = "1cm") %>%
+  column_spec(5:6, width = "1cm") 
+```

### 2020-05-19 manuscript.Rmd
* minor revisions * completed refs

@@
 
-Insufficient physical activity is a leading cause of disease and disability worldwide.[@who2009] For the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic background are disproportionately affected by physical inactivity-related diseases.
-Increasing the physical activity levels of the population is therefore high on the public health agenda: it not only has the potential to improve quality of life, reduce mortality rates and alleviate the strain on health and social care services, but also reduce the gap in health inequalities.
+Insufficient physical activity is a leading cause of disease and disability worldwide.[@who2009] For the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic backgrounds are both disproportionately likely to be inactive [@rowe2000sport; @farrell2014] and be affected by physical inactivity-related diseases.
+Increasing the physical activity levels of the population is therefore high on the public health agenda: it not only has the potential to improve quality of life, reduce mortality rates and alleviate the strain on health and social care services, but also reduce the gap in health inequalities [@vilhelmsson2018].
 
-However, designing effective public health interventions that increase population physical activity is considerable challenge.[@ref] Implementing such interventions in a way that does not increase health inequalities might even be more difficult, as studies have shown that programmes to increase physical activity often fail to reach deprived communities and those most in need.[@white2009; @rideout2018]
+However, designing effective public health interventions that increase population physical activity is considerable challenge.[@reis2016] Implementing such interventions in a way that does not increase health inequalities might even be more difficult, as studies have shown that programmes to increase physical activity often fail to reach deprived communities and those most in need.[@white2009; @rideout2018]
 
 parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professionals with respect to both.
-Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2004; @wiltshire2018]
+Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@haake2018pr; @reece2019] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2013] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2019changes; @wiltshire2018]
 
@@
 It might therefore be the case that parkrun events are primarily located in areas that are relatively well-off anyway, while people living in more deprived communities may not have the same opportunities to participate.
-In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@sportengland2018]
+In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@reuters2018]
 
@@
 
-  1) For all 32,844 LSOAs, we retrieved geographic locations, defined by the coordinates of its population-weighted centroid, 2017 total population estimates, and the 2015 Index of Multiple Deprivation (IMD) from the Office for National Statistics [@ons2018; @ref; @ref].
+  1) For all 32,844 LSOAs, we retrieved geographic locations, defined by the coordinates of its population-weighted centroid, 2017 total population estimates, and the 2015 Index of Multiple Deprivation (IMD) from the Office for National Statistics.[@ons2011; @ons2015; ons2017]
 
-  2) We included all 465 public parkrun events which were in operation in England by December 12th, 2018 - on this date, Sport England announced their plan to provide funding for setting up 200 additional parkrun events across England.[@ref] The locations of the events were manually scrapped from the parkrun UK website.[@ref]
+  2) We included all 465 public parkrun events which were in operation in England by December 12th, 2018 - on this date, Sport England announced their plan to provide funding for setting up 200 additional parkrun events across England.[@parkrunGroup] The locations of the events were manually scrapped from the parkrun UK website.[@parkrunweb]
 
-  3) The locations of public green spaces in England were retrieved from an open dataset of Ordnance Survey.[9] Even though parkrun events do not always require a 5km loop (some courses involve running back and forth) and are held in various settings and terrains, not all green spaces are suitable to host running events.
+  3) The locations of public green spaces in England were retrieved from an open dataset of Ordnance Survey.[@ordnance] Even though parkrun events do not always require a 5km loop (some courses involve running back and forth) and are held in various settings and terrains, not all green spaces are suitable to host running events.
 After evaluating existing parkrun event courses, we decided to consider all public parks, gardens, and playing fields with an area of 0.1 km$^2$ (e.g.
@@
 
-  In order to identify the optimal new locations for setting up 200 new parkun events consecutively, the selection procedure is repeated 200 times.
+  In order to identify the optimal new locations for setting up 200 new parkrun events consecutively, the selection procedure is repeated 200 times.
 At each step, the single best candidate green space location is selected, added to the set of established parkrun events $E$, and removed from the set of available green spaces $C$.
@@
 
-  All data and the R source code that were used to generate the results of this study are provided on an open repository.[@ref]
+  All data and the R source code that were used to generate the results of this study are provided on an open repository.[@githubrepo]
 
@@
 There was a negative relationship between IMD and the distance to the nearest parkrun event: the (population weighted) Pearson and Spearman correlation coefficients were -0.15 and -0.18, indicating a small effect.
-This means, more deprived LSOAs tended to have shorter distances to the nearest parkun event, i.e.
+This means, more deprived LSOAs tended to have shorter distances to the nearest parkrun event, i.e.
 better access.
@@
 
-We estimated that setting up new parkrun events in those 200 green spaces would improve access for around 16.5 million people (30% of th population) from 9,854 LSOAs.
+We estimated that setting up new parkrun events in those 200 green spaces would improve access for around 16.5 million people (30% of the population) from 9,854 LSOAs.
 For those, the distance to the nearest event would, on average, be reduced by 4.09 km (SD = 3.97).
@@
 
-Overall, geographic access to parkrun in England is good. On 12th December 2018 
-The median distance to the nearest parkrun event was 3.39 km and more than two thirds of the population lived within 5km (the parkrun distance) of a parkrun event.
+  Overall, geographic access to parkrun in England is good: the median distance to the nearest parkrun event was 3.39 km and more than two thirds of the population lived within 5km (the parkrun distance) of a parkrun event.
 Contrary to our expectation, we did not find that access was better for people living in more affluent areas.
@@
   The main finding, that geographical access to parkrun events is better in more deprived communities, is surprising.
-Parkrun events are set up by volunteers, based on demand, not need.
+parkrun events are set up by volunteers, based on demand, not need.
 Studies have shown that the level of physical activity, and the availability of physical activity facilities generally declines with the level of deprivation.
@@
 @hillsdon2007] parkrun events, in contrast, seem to be often held in or near deprived areas, and are free to attend, giving anyone equal access, irrespective of their socio-economic background.
-Nevertheless, in a previous analysis, we found that participation in parkrun has a strong socioeconomic gradient with considerably higher participation rates in more affluent areas: about a third of all participants came from LSOAs in the least deprived quintile, while only 7% came from the most deprived quintile.[@smithschneider; @schneidersmith] This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been sufficient to engage deprived communities.
+Nevertheless, in a previous analysis, we found that participation in parkrun has a strong socioeconomic gradient with considerably higher participation rates in more affluent areas: about a third of all participants came from LSOAs in the least deprived quintile, while only 7% came from the most deprived quintile.[@smithschneider; @schneidersmith] This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been sufficient to engage deprived communities.[@white2009]
 
@@
 Most importantly geographic access is not measured as travel distance, or travel time.
-In some cases, for example where natural barriers like lakes or rivers block routes, the actual distance travelled may be far in excess of the distance as the crow flies.
+In some cases, for example where natural barriers like lakes or rivers block routes, the actual distance travelled may be far in excess of the distance as the crow flies.[@boscoe2012]
 
@@
 We did not investigate what other factors independently explain access more generally.
-It should be noted, however, that a contributing factor for the negative relationship between IMD and access is likely to be population density: deprived areas cluster in urban areas, where also most parkun events take place.
+It should be noted, however, that a contributing factor for the negative relationship between IMD and access is likely to be population density: deprived areas cluster in urban areas, where also most parkrun events take place.
 Rural areas, on the other hand, may therefore have worse geographical access as the crow flies.
@@
 
-  In England, parkrun has been hugely successful in engaging large numbers of people in physical activity.
-The inclusive culture of their events, and the almost universal availability of parkrun events throughout the country provide a learning opportunity to explore socio-spatial determinants influencing physical activity behaviour on a national scale.
-Our study contributes to the limited research in this area and identifies possible leads for further investigations.
+  In England, parkrun has been hugely successful in engaging large numbers of people in physical activity.[@reece2019] The inclusive culture of their events, and the almost universal availability of parkrun events throughout the country provide a learning opportunity to explore socio-spatial determinants influencing physical activity behaviour on a national scale.[@wiltshire2018] Our study contributes to the limited research in this area and identifies possible leads for further investigations.
 Studying barriers to participation in parkrun, other than geographic access, is likely to improve our understanding of the reasons why physical activity levels are lower in more deprived areas and may help to design more effective public health interventions to increase levels of physical activity in the population.
@@
 
-## References
-
 <div id="refs"></div>
@@
 ```{r echo=FALSE}
-
   kable(read.csv("./output/map_legend.csv"),

### 2020-05-19 manuscript.Rmd

@@
 
-*$^{1}$School of Health and Related Research, University of Sheffield, Sheffield, UK.; $^{2}$Advanced Wellbeing Research Centre, Sheffield Hallam University, Sheffield, UK.*
+*$^{1}$School of Health and Related Research, University of Sheffield, Sheffield, UK.;    
+$^{2}$Advanced Wellbeing Research Centre, Sheffield Hallam University, Sheffield, UK.*
+
+Corresponding author: Paul Schneider (p.schneider@sheffield.ac.uk)
+
+Target Journal: 1) Public Health (https://www.journals.elsevier.com/public-health), 2) BMJ Open, 3) Wellcome Open Research
+
+\newpage
+
+## Abstract
+
+### Objectives
+
+To evaluate geographical access to parkrun in England, and to identify 200 locations for future events to maximize geographic access to parkrun.
+
+### Study design
+
+This study is a cross-sectional ecological analysis of the socio-economic disparities in geographic access to parkrun events in England at the end of 2018.
+
+### Methods
+
+We combined geo-location data on all English Lower Layer Super Output Areas and parkrun events to calculate geodesic distances to the nearest event for over 32,000 communities in England.
+We use this measure of geographic access to summarise the relationship between access and socio-economic deprivation, measured using the Index of Multiple Deprivation.
+We then used geographic coordinates of public green spaces in England to conduct a simple location-allocation analysis to identify 200 locations for future parkrun events that would maximize access.
+
+### Results
+
+Geographical access to parkrun in England is good, over two thirds of the population live within 5km of a parkrun.
+It is also very equitable, access is slightly better in more deprived communities.
+Optimally locating an additional 200 events would improve access considerably, approximately 82% of the English population would be within 5km of a parkrun event.
+
+### Conclusions
+
+Access to parkrun is good, and equitable.
+Creating new parkrun events will improve geographic access to parkrun, but a deeper understanding of the barriers to participation in deprived communities is necessary in order to develop effective strategies to improve engagement.
+
+\newpage
 
@@
 
-Insufficient physical activity is a leading cause of disease and disability worldwide.[@who2009] For the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic backgrounds are both disproportionately likely to be inactive [@rowe2000sport; @farrell2014] and be affected by physical inactivity-related diseases.
+Insufficient physical activity is one of the leading causes of disease and disability worldwide.[@who2009] In the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic backgrounds are both disproportionately likely to be inactive [@rowe2000sport; @farrell2014] and be affected by physical inactivity-related diseases [@petrovic2018contribution].
 Increasing the physical activity levels of the population is therefore high on the public health agenda: it not only has the potential to improve quality of life, reduce mortality rates and alleviate the strain on health and social care services, but also reduce the gap in health inequalities [@vilhelmsson2018].
@@
 
-Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely driven by *demand*, not *need*.
-It might therefore be the case that parkrun events are primarily located in areas that are relatively well-off anyway, while people living in more deprived communities may not have the same opportunities to participate.
+Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely grassroots, driven by *demand* rather than *need*.
+It might therefore be the case that parkrun events are primarily located in areas that are relatively affluent, while people living in more deprived communities may not have the same opportunities to participate.
 In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@reuters2018]
@@
 The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
-Secondly, to identify 200 optimal locations for future events in order geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
 Secondly, to identify 200 optimal locations for future events in order to improve geographic access, in particular for deprived communities.
@@
 We assessed the relationship between access, defined as the distance (as the crow flies) to the nearest parkrun event, and socio-economic deprivation, measured using the Index of Multiple Deprivation.
-In addition, we used information on public green spaces in England to conduct a simple location-allocation analysis, in order to identify 200 locations for future parkrun events, that maximise access for the population.
+In addition, we used information on public green spaces in England to conduct a simple location-allocation analysis, in order to identify 200 locations for future parkrun events that maximise access for the population.
 
@@
 
-  3) The locations of public green spaces in England were retrieved from an open dataset of Ordnance Survey.[@ordnance] Even though parkrun events do not always require a 5km loop (some courses involve running back and forth) and are held in various settings and terrains, not all green spaces are suitable to host running events.
-After evaluating existing parkrun event courses, we decided to consider all public parks, gardens, and playing fields with an area of 0.1 km$^2$ (e.g.
-316m x 316m) or more potentially suitable for hosting events (n=2,842).
+  3) The locations of public green spaces in England were retrieved from an open dataset of Ordnance Survey.[@ordnance] parkrun events are held in various settings and terrains, and do not always require a single 5km loop - some events have courses that involve running back and forth.
+After evaluating existing parkrun event courses, we decided to consider all public parks, gardens, and playing fields in England with an area of 0.1 km$^2$ or more potentially suitable for hosting events (n=2,842).
 
@@
 
-### Association between Deprivation and Access
+### Association between deprivation and access
 
-There was a negative relationship between IMD and the distance to the nearest parkrun event: the (population weighted) Pearson and Spearman correlation coefficients were -0.15 and -0.18, indicating a small effect.
-This means, more deprived LSOAs tended to have shorter distances to the nearest parkrun event, i.e.
-better access.
+There was a negative relationship between IMD and the distance to the nearest parkrun event: the (population weighted) Pearson and Spearman correlation coefficients were -0.15 and -0.18, indicating a small negative correlation.
+This means that more deprived LSOAs tended to have shorter distances to the nearest parkrun event, i.e.
+better geographical access.
 
@@
 We estimated that setting up new parkrun events in those 200 green spaces would improve access for around 16.5 million people (30% of the population) from 9,854 LSOAs.
-For those, the distance to the nearest event would, on average, be reduced by 4.09 km (SD = 3.97).
+For these people, the distance to the nearest event would, on average, be reduced by 4.09 km (SD = 3.97).
 Overall, it would reduce the average and median distance to the nearest parkrun event from 4.65 and 3.39 km to 3.43 and 2.59 km.
-The percentage of people who live within 5km of a parkrun would increase from 69% to 81.9%.
+The percentage of people who live within 5km of a parkrun would increase from 69% to 82%.
 Figure 2 shows the distribution of distances between LSOAs and their nearest parkrun events before and after the introduction of the 200 new optimal event locations at the identified optimal green spaces.
@@
 
-  Overall, geographic access to parkrun in England is good: the median distance to the nearest parkrun event was 3.39 km and more than two thirds of the population lived within 5km (the parkrun distance) of a parkrun event.
+  Overall, geographic access to parkrun in England is good: in December 2018 the median distance to the nearest parkrun event was 3.39 km and more than two thirds of the population lived within 5km (the parkrun distance) of a parkrun event.
 Contrary to our expectation, we did not find that access was better for people living in more affluent areas.
@@
 parkrun events are set up by volunteers, based on demand, not need.
-Studies have shown that the level of physical activity, and the availability of physical activity facilities generally declines with the level of deprivation.
-Opportunities for physical activity are often lacking in areas in most need.[e.g.
+Studies have shown that the level of physical activity, and the availability of physical activity facilities generally declines with the level of deprivation.[@farrell2014socioeconomic] Opportunities for physical activity are often lacking in areas in most need.[e.g.
 @hillsdon2007] parkrun events, in contrast, seem to be often held in or near deprived areas, and are free to attend, giving anyone equal access, irrespective of their socio-economic background.
@@
 Given that participation rates are generally lower in deprived areas, improving access alone is unlikely to significantly reduce inequalities in participation and/or levels of physical activity.
-A deeper understanding is needed of the barriers to participation in deprived communities if strategies are going to be created to improve engagement.
+A deeper understanding is needed of the barriers to participation in deprived communities if effective strategies are going to be created to improve engagement.
+
+## Additional Information
+
+### Grant Information
+P.S, R.S & T.B are joint funded by the Wellcome Trust Doctoral Training Centre in Public Health Economics and Decision Science [108903] and the University of Sheffield.
+
+### Competing Interests
+P.S, R.S, T.B & A.Br have no competing interests.
+S.H. is chair, A.Bu. and H.Q are deputy chairs, and E.G. is a member of the parkrun research board.
+
+### Software & Data Availability
+All data and source code are available at: TBC (<add final GitHub repo>).
+
+### Author contributions
+
+Paul Schneider; Roles: Conceptualization, Data Curation, Formal Analysis, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing
+
+Robert Smith; Roles: Conceptualization, Data Curation, Formal Analysis, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing
+
+Alice Bullas; Roles: Writing – Review & Editing
+
+Steve Haake; Writing – Review & Editing
+
+Helen Quirk; Roles: Writing – Review & Editing
+
+Thomas Bayley; Roles:  Methodology
+
+Alan Brennan; Roles: Supervision, Methodology, Writing – Review & Editing
+
+Elizabeth Goyder; Roles: Supervision, Conceptualization, Writing – Review & Editing
+
+\newpage
+
+## References
 

### 2020-05-26 src/src4_analysis.R
added scales number format.

@@ -100,6 +100,7 @@
       geom_density(aes(mn_dstn,fill="Current access to parkrun events (December 12th, 2018)"),col="black",alpha=0.3) +
       scale_fill_manual(values = c( "Current access to parkrun events (December 12th, 2018)" = "red","After 200 new parkrun events are set up"="lightgreen"),name="") +
       scale_x_continuous(label=label_number(suffix = " km"),limits = c(0,25),
+                         scales::number_format(accuracy = 1),
                          name = "Distance to the nearest parkrun event") +
       theme_minimal()+
       theme(legend.position = "bottom")

### 2020-05-26 src/src4_analysis.R
added scales to label

@@ -99,8 +99,8 @@
       geom_density(aes(mn_dstn_new,fill="After 200 new parkrun events are set up"),col="black",alpha=0.5) +
       geom_density(aes(mn_dstn,fill="Current access to parkrun events (December 12th, 2018)"),col="black",alpha=0.3) +
       scale_fill_manual(values = c( "Current access to parkrun events (December 12th, 2018)" = "red","After 200 new parkrun events are set up"="lightgreen"),name="") +
-      scale_x_continuous(label=label_number(suffix = " km"),limits = c(0,25),
-                         scales::number_format(accuracy = 1),
+      scale_x_continuous(label=scales::number_format(accuracy = 1,suffix = " km"),
+                         limits = c(0,25),
                          name = "Distance to the nearest parkrun event") +
       theme_minimal()+
       theme(legend.position = "bottom")

### 2020-05-27 manuscript.Rmd
prefinal draft

@@
 
-  Schneider PP$^{*,1}$, Smith RA$^{1}$, Bullas AM$^{2}$, Quirk H$^{2}$, Bayley T$^{1}$, Haake SSJ$^{2}$, Brennan A$^{1}$, Goyder E$^{1}$
+  Schneider PP$^{*,1}$, Smith RA$^{1}$, Bullas AM$^{3}$, Quirk H$^{1}$, Bayley T$^{1}$, Haake SSJ$^{2}$, Brennan A$^{1}$, Goyder E$^{1}$
 
 *$^{1}$School of Health and Related Research, University of Sheffield, Sheffield, UK.;    
-$^{2}$Advanced Wellbeing Research Centre, Sheffield Hallam University, Sheffield, UK.*
+$^{2}$Advanced Wellbeing Research Centre, Sheffield Hallam University, Sheffield, UK.
+$^{3}$The Centre for Sports Engineering Research, Sheffield Hallam University, Sheffield S9 3TU, UK*
 
@@
 
-Target Journal: 1) Public Health (https://www.journals.elsevier.com/public-health), 2) BMJ Open, 3) Wellcome Open Research
+Target Journal: 1) Public Health (https://www.journals.elsevier.com/public-health)
 
@@
 
-To evaluate geographical access to parkrun in England, and to identify 200 locations for future events to maximize geographic access to parkrun.
+To evaluate geographic access to free weekly outdoor physical activity events (“parkrun”) in England, with a particular focus on deprived communities, and to identify 200 locations for future events to maximize geographic access to parkrun.
 
@@
 We use this measure of geographic access to summarise the relationship between access and socio-economic deprivation, measured using the Index of Multiple Deprivation.
-We then used geographic coordinates of public green spaces in England to conduct a simple location-allocation analysis to identify 200 locations for future parkrun events that would maximize access.
+We then used geographic coordinates of public green spaces in England to conduct a simple location-allocation analysis to identify 200 locations for future event locations that would maximize access.
 
@@
 
-Geographical access to parkrun in England is good, over two thirds of the population live within 5km of a parkrun.
-It is also very equitable, access is slightly better in more deprived communities.
-Optimally locating an additional 200 events would improve access considerably, approximately 82% of the English population would be within 5km of a parkrun event.
+In England 69% of the population live within 5 km of one of the 465 parkrun events.
+There is a small negative correlation between distance and deprivation, indicating that access is slightly better in more socioeconomically deprived areas.
+Setting up an additional 200 events in optimal locations would improve access: the average distance to the nearest parkrun event would improve by 1.22 km, from 4.65 km to 3.43 km, and approximately 82% of the English population would live within 5 km of a parkrun event.
 
@@
 
-Access to parkrun is good, and equitable.
-Creating new parkrun events will improve geographic access to parkrun, but a deeper understanding of the barriers to participation in deprived communities is necessary in order to develop effective strategies to improve engagement.
+Over two thirds of the English population live within 5 km of a parkrun event, and contrary to our expectation, we find that geographic access is slightly better for those living in more deprived communities.
+Creating additional events may improve geographic access, but effective strategies will still be needed to increase engagement in new and existing events by those living in socioeconomically deprived areas.
 
@@
 
-However, designing effective public health interventions that increase population physical activity is considerable challenge.[@reis2016] Implementing such interventions in a way that does not increase health inequalities might even be more difficult, as studies have shown that programmes to increase physical activity often fail to reach deprived communities and those most in need.[@white2009; @rideout2018]
+However, designing effective public health interventions that increase population physical activity is considerable challenge.[@reis2016] Implementing such interventions in a way that does not increase health inequalities might even be more difficult.
+Studies have shown that programmes to increase physical activity often fail to reach deprived communities and those most in need, suggesting access to physical activity facilities shows spatial and social inequality.[@white2009; @rideout2018] In this regard, parkrun, an international movement which organises free weekly 5 km running and walking events in public spaces, might provide valuable lessons.
 
-parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professionals with respect to both.
-Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@haake2018pr; @reece2019] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2013] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2019changes; @wiltshire2018]
+Since its founding in 2004, as a small event in London with 13 participants, parkrun has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@haake2018pr; @reece2019] The volunteer-led events are often characterised as accessible and inclusive.[@stevinson2013; @quirk2019] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2019changes; @wiltshire2018]
 
 Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely grassroots, driven by *demand* rather than *need*.
-It might therefore be the case that parkrun events are primarily located in areas that are relatively affluent, while people living in more deprived communities may not have the same opportunities to participate.
+It might therefore be the case that parkrun events are primarily located in areas that are less deprived, while people living in more deprived communities may not have the same opportunities to participate.
 In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@reuters2018]
 
-The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
+The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
 Secondly, to identify 200 optimal locations for future events in order to improve geographic access, in particular for deprived communities.
@@
 
-  2) We included all 465 public parkrun events which were in operation in England by December 12th, 2018 - on this date, Sport England announced their plan to provide funding for setting up 200 additional parkrun events across England.[@parkrunGroup] The locations of the events were manually scrapped from the parkrun UK website.[@parkrunweb]
+  2) We included all 465 public parkrun events which were in operation in England by December 12th, 2018 - on this date, Sport England announced their plan to provide funding for setting up 200 additional parkrun events across England.[@parkrunGroup] The locations of the events were obtained from the parkrun UK website.[@parkrunweb]
 
-  3) The locations of public green spaces in England were retrieved from an open dataset of Ordnance Survey.[@ordnance] parkrun events are held in various settings and terrains, and do not always require a single 5km loop - some events have courses that involve running back and forth.
+  3) The locations of public green spaces in England were retrieved from an open dataset of Ordnance Survey.[@ordnance] parkrun events are held in various settings and terrains, and do not always require a single 5 km loop - some events have courses that involve running a combination of shorter loops..
 After evaluating existing parkrun event courses, we decided to consider all public parks, gardens, and playing fields in England with an area of 0.1 km$^2$ or more potentially suitable for hosting events (n=2,842).
@@
 
-  Access to parkrun was measured in terms of the distance as the crow flies (in km) to the nearest event.
+  Access to parkrun was defined as the geodesic distance (as the crow flies) from LSOA’s population-weighted centroid to its nearest event.
 For each of the 32,844 LSOAs, we computed the geodesic distances between its population-weighted centroid and all 465 parkun events that were in operation on December 12th, 2018, and then selected the shortest distance.
@@
 population density or demographics), were not taken into account, because we did not aim to assess to what extent deprivation independently 'explained' access.
-Rather, we sought to evaluate whether or not people living in deprived areas have better or worse access, given the actual circumstances.
+Rather, we sought to evaluate whether or not people living in deprived areas have better or worse geographic access, under the actual circumstances.
 
@@
 We then assessed the association between the IMD and the distance to the nearest parkrun event on the LSOA level.
-Our hypothesis was that more deprived areas had worse access, i.e.
-longer distances to the nearest parkrun event than more affluent areas.
+Our hypothesis was that more socioeconomically deprived areas had worse access, i.e.
+longer geodesic distances to the nearest parkrun event than less deprived areas.
 Pearson and Spearman correlation coefficients were computed using the LSOAs' total population as weights.
-We also conducted a stratified analysis, for which we grouped LSOAs into IMD quintiles (most, more, median, less, least deprived) and assessed access to parkrun events in each strata.
+We also conducted a stratified analysis, for which we grouped LSOAs into IMD quintiles (most, more, median, less, least deprived) and assessed access to parkrun events in each stratum.
 
@@
 
-  We conducted a location-allocation analysis to solve the following problem: parkrun UK received funding to start 200 additional parkrun events.
-There are 2,842 public green spaces in England in which new events could be set up.
-Which 200 locations should be selected, in order to maximise access for the greatest number of people?
+  We conducted a location-allocation analysis to solve the following problem.
+parkrun UK received funding to start 200 additional parkrun events; there are 2,842 public green spaces in England in which new events could be set up – which 200 locations should be selected, in order to maximise access for the greatest number of people?
 
@@
 Firstly, for each green space, we evaluate how setting up a parkrun event would affect the sum of distances, given the locations of all existing events (i.e.
-for how many LSOAs would this green space then be the nearest parkrun event, and by how much would it decrease the respective distances).
+for how many LSOAs this green space would be the nearest parkrun event, and by how much it would decrease the respective distances).
 Secondly, the green space with the greatest effect is selected and added to the set of existing parkrun events.
@@
 
-  As of 12th December 2018, approximately 7%, 69%, and 91% of the English population lived within 1, 5, and 10km distance of a parkrun event.
-Only 578,043 people (1% of the English population) lived more than 20km from an event.
+  As of 12th December 2018, approximately 7%, 69%, and 91% of the English population lived within 1, 5, and 10 km distance of a parkrun event.
+Only 578,043 people (1% of the English population) lived more than 20 km from an event.
 The average (SD) and median (IQR) distance to the nearest parkrun event were 4.65 (4.22) and 3.39 (1.99-5.83).
-The worst access was observed for 2,259 people living on the Isles of Scilly, who live about 76.44km (as the crow flies) away from the next parkrun event on the mainland.
+The largest distance was observed for the 2,259 people living on the Isles of Scilly, who live about 76 km away from the next parkrun event on the mainland.
+On average, each parkrun event is the closest event for 71 LSOAs (43), with a combined population of 119,612 (74,290).
 Further descriptive statistics are provided in table 1.
@@
 This means that more deprived LSOAs tended to have shorter distances to the nearest parkrun event, i.e.
-better geographical access.
+better geographic access.
 
-The distributional analysis of distances by IMD quintile showed that people living in the 20% most deprived LSOAs  had the best geographic access, with an average and median distance to the nearest parkrun event of 3.51 and 2.79 km.
-Depending on the metric, the worst access was observed for LSOAs in the middle (mean distance = 3.36km) or the less deprived group (median distance = 3.93km).
-Further results of the distributional analysis are shown in Table 2 in the following section.
+The analysis of distances by IMD quintile in Table 2 shows that people living in the 20% most deprived LSOAs had the best geographic access, with a mean and median distance to the nearest parkrun event of 3.51 and 2.79 km.
+Depending on the metric, the worst access was observed for LSOAs in the middle (mean distance = 3.36 km) or the less deprived group (median distance = 3.93 km).
+Further results of the distributional analysis are provided in Table 2.
 
@@
 
-Figure 1 shows the parkrun events (black circles) that existed on 12th December 2018 alongside recommendations for 200 additional event locations (red triangles), which maximise overall access to parkrun for the greatest number of people.
+Figure 1 shows the parkrun events (black circles) that existed on 12th December 2018 alongside recommendations for 200 additional event locations (red triangles), which minimise the sum of the population weighted geodesic distances from the LSOA centroids, i.e.
+maximise overall access to parkrun for the greatest number of people.
 The numbers correspond to the rank, where 1 is the location which would improve access the most.
@@
 ![](./output/map_static.jpeg)
-**Figure 1: Map of England showing current parkrun events (blue circles) and recommended new event locations (red triangles) ranked in descending order of estimated effect on overall population-weighted access.
-Information on all 200 identified optimal green space locations are provided in the appendix.**
+**Figure 1**: Map of England showing current parkrun events (blue circles) and recommended new event locations (red triangles) ranked in descending order of estimated effect on overall population-weighted access.
+Information on all 200 identified optimal green space locations are provided in the appendix.
 
@@
 Overall, it would reduce the average and median distance to the nearest parkrun event from 4.65 and 3.39 km to 3.43 and 2.59 km.
-The percentage of people who live within 5km of a parkrun would increase from 69% to 82%.
-Figure 2 shows the distribution of distances between LSOAs and their nearest parkrun events before and after the introduction of the 200 new optimal event locations at the identified optimal green spaces.
+The percentage of people who live within 5 km of a parkrun would increase from 69% to 82%.
 
-![Figure 2](./output/hist.png)
-
-**Figure 2: Density plot showing the distance, as the crow flies, to parkrun events before and after the creation of 200 new events in identified green-spaces. **
-
-Table 2 shows further results for the effects on LSOAs across IMD quintiles.
-Overall, setting up 200 new events in the recommended green spaces would amplify the negative association between IMD and geographic access: the population-weighted Pearson and Spearman correlation coefficients changed from -0.15 and -0.18 before, to -0.20 and -0.23 afterwards, indicating that improvements in access to parkrun events were greater for more deprived LSOAs as compared to less deprived LSOAs.
+The distributional analysis in Table 2 shows, for each IMD quintile, geographic access under the current situation (12th December 2018) and after the creation of 200 new events.
+Overall, setting up 200 new events in the recommended green spaces would amplify the negative socioeconomic gradient in geographic access.
+The population-weighted Pearson and Spearman correlation coefficients changed from -0.15 and -0.18 before, to -0.20 and -0.23 afterwards, indicating that improvements in access to parkrun events were greater for more deprived LSOAs as compared to less deprived LSOAs.
 Nevertheless, the distributional analysis showed that the improvement in access was smallest for LSOAs in the most deprived quintile.
@@
 
-  Overall, geographic access to parkrun in England is good: in December 2018 the median distance to the nearest parkrun event was 3.39 km and more than two thirds of the population lived within 5km (the parkrun distance) of a parkrun event.
-Contrary to our expectation, we did not find that access was better for people living in more affluent areas.
-In fact, those living in the most deprived areas had the best geographic access to parkrun.
-It is rare in public health for inequalities to exist in this direction.
+As of 12 December 2018, the median distance to the nearest parkrun event was 3.39 km and more than two thirds of the English population lived within 5 km (the parkrun distance) of a parkrun event.
+Contrary to our expectation, we did not find that access was better for people living in less deprived areas.
+In fact, those living in the most deprived areas had the best geographic access to parkrun – it is rare in public health for inequalities to exist in this direction.
 
-  Our analysis has shown that setting up 200 new events in the recommended (optimal) green spaces in England would considerably improve access - for example, the percentage of English residents who live within 5km of a parkrun would increase to 82%.
-Moreover, the recommended expansion of parkrun would improve the geographic access for the most deprived communities more than the access of those living in more affluent areas.
+  Our analysis has shown that setting up 200 new events in the recommended (optimal) green spaces in England would reduce the average distance to the nearest parkrun event by 1.22 km, increasing the percentage of English residents who live within 5 km of a parkrun to 82%.
+Moreover, the recommended expansion of parkrun would improve the geographic access for the most deprived areas more than the access of those living in more affluent areas.
 
-  The main finding, that geographical access to parkrun events is better in more deprived communities, is surprising.
+  The main finding, that geographic access to parkrun events is better in more deprived communities, is surprising.
 parkrun events are set up by volunteers, based on demand, not need.
@@
 @hillsdon2007] parkrun events, in contrast, seem to be often held in or near deprived areas, and are free to attend, giving anyone equal access, irrespective of their socio-economic background.
-Nevertheless, in a previous analysis, we found that participation in parkrun has a strong socioeconomic gradient with considerably higher participation rates in more affluent areas: about a third of all participants came from LSOAs in the least deprived quintile, while only 7% came from the most deprived quintile.[@smithschneider; @schneidersmith] This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been sufficient to engage deprived communities.[@white2009]
+Nevertheless, in a previous analysis, we found that participation in parkrun has a strong socioeconomic gradient with considerably higher participation rates in less deprived areas: about a third of all participants came from LSOAs in the least deprived quintile, while only 7% came from the most deprived quintile.[@smithschneider; @schneidersmith] This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been sufficient to engage people living in deprived communities.[@white2009] This means, creating additional events in optimal locations could improve overall geographic access further, yet effective strategies will still be needed to increase equity in engagement in new and current events.
 
@@
 Secondly the analysis makes use of large and rich datasets, with over 30,000 LSOA and over 400 existing parkrun events it is unlikely that individual outliers are affecting the results.
+The almost universal availability of parkrun events throughout the country provides a learning opportunity to explore socio-spatial determinants influencing physical activity behaviour on a national scale.[@wiltshire2018] Our study contributes to the limited research in this area and identifies possible leads for further investigation.
 
   However, there are also limitations.
-Most importantly geographic access is not measured as travel distance, or travel time.
-In some cases, for example where natural barriers like lakes or rivers block routes, the actual distance travelled may be far in excess of the distance as the crow flies.[@boscoe2012]
+Most importantly, geographic access is not measured as travel distance, or travel time, but as geodesic distance.
+In some cases, for example where natural barriers like hills, lakes or rivers block routes, the actual distance travelled may be far in excess of the geodesic distance.[@boscoe2012]
 
@@
 It should be noted, however, that a contributing factor for the negative relationship between IMD and access is likely to be population density: deprived areas cluster in urban areas, where also most parkrun events take place.
-Rural areas, on the other hand, may therefore have worse geographical access as the crow flies.
+Rural areas, on the other hand, may therefore have worse geographic access.
 Further studies are required to better understand wider determinants of access to parkrun and/or physical activity facilities more generally.
 
-  In England, parkrun has been hugely successful in engaging large numbers of people in physical activity.[@reece2019] The inclusive culture of their events, and the almost universal availability of parkrun events throughout the country provide a learning opportunity to explore socio-spatial determinants influencing physical activity behaviour on a national scale.[@wiltshire2018] Our study contributes to the limited research in this area and identifies possible leads for further investigations.
 Studying barriers to participation in parkrun, other than geographic access, is likely to improve our understanding of the reasons why physical activity levels are lower in more deprived areas and may help to design more effective public health interventions to increase levels of physical activity in the population.
@@
 
-  In England, geographic access to parkrun events is good, and can be further improved with the creation of 200 new events.
-Contrary to our expectation, we find that geographic access is better for those living in more deprived communities.
-Given that participation rates are generally lower in deprived areas, improving access alone is unlikely to significantly reduce inequalities in participation and/or levels of physical activity.
-A deeper understanding is needed of the barriers to participation in deprived communities if effective strategies are going to be created to improve engagement.
+  In England in December 2018, 69% of the population lived within 5 km of a parkrun event.
+Creating 200 new events in the recommended (optimal) green spaces would further improve access, increasing this to 82%.
+Contrary to our expectation, we find that geographic access is slightly better for those living in more deprived communities.
+Given that participation rates are generally lower in deprived areas, improving access alone seems unlikely to significantly reduce inequalities in participation and physical activity.
+In order to design more effective strategies to improve engagement from deprived communities, a deeper understanding of  the barriers to taking part in mass participation physical activity events is needed.
 

### 2020-05-27 src/src2_data_proc.R
minor fix

@@ -135,7 +135,7 @@
       greens_sp = raster::shapefile("./input/greenspaces_data/trimmed_greenspaces")
   
 # CLEAN UP
-    rm(list=c("lsoa_min_dist_event","srvd_lsoa","srvd_pop","MIN_PARK_AREA","greens_in_england","lsoa_imd","lsoa_pop","lsoa_cntrds"))
+    rm(list=c("lsoa_min_dist_event","srvd_lsoa","srvd_pop","MIN_PARK_AREA","greens_in_england","lsoa_imd","lsoa_pop"))
 
 # SAVE the created data sets (saves time for later use)
     save(list=c("dist_M_full","event_sp","greens_sp","lsoa_min_dist","lsoa_sp"),

### 2020-05-29 manuscript.Rmd
prefinal revision fpr submission

@@
 csl: vancouver.csl
+linestretch: 1.3
+fontsize: 12pt
+indent: false
 ---
@@
 
-*$^{1}$School of Health and Related Research, University of Sheffield, Sheffield, UK.;    
+*$^{1}$School of Health and Related Research, University of Sheffield, Sheffield, UK.   
 $^{2}$Advanced Wellbeing Research Centre, Sheffield Hallam University, Sheffield, UK.   
-$^{3}$The Centre for Sports Engineering Research, Sheffield Hallam University, Sheffield S9 3TU, UK*
+$^{3}$The Centre for Sports Engineering Research, Sheffield Hallam University, Sheffield S9 3TU, UK.*
 
@@
 
-Target Journal: 1) Public Health (https://www.journals.elsevier.com/public-health)
+Target Journal: Public Health (https://www.journals.elsevier.com/public-health)
 
 \newpage
+
+# Multiple deprivation and geographic distance to community sport events — achieving equitable access to parkrun in England
 
@@
 
-This study is a cross-sectional ecological analysis of the socio-economic disparities in geographic access to parkrun events in England at the end of 2018.
+This study is a cross-sectional ecological analysis of the socioeconomic disparities in geographic access to parkrun events in England at the end of 2018.
 
@@
 We combined geo-location data on all English Lower Layer Super Output Areas and parkrun events to calculate geodesic distances to the nearest event for over 32,000 communities in England.
-We use this measure of geographic access to summarise the relationship between access and socio-economic deprivation, measured using the Index of Multiple Deprivation.
+We use this measure of geographic access to summarise the relationship between access and socioeconomic deprivation, measured using the Index of Multiple Deprivation.
 We then used geographic coordinates of public green spaces in England to conduct a simple location-allocation analysis to identify 200 locations for future event locations that would maximize access.
@@
 
-Insufficient physical activity is one of the leading causes of disease and disability worldwide.[@who2009] In the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic backgrounds are both disproportionately likely to be inactive [@rowe2000sport; @farrell2014] and be affected by physical inactivity-related diseases [@petrovic2018contribution].
+Insufficient physical activity is one of the leading causes of disease and disability worldwide [@who2009].
+In the UK, around one in six deaths is attributable to low levels of physical activity [@lee2012].
+It is also a major contributor to health inequalities, as people from low socioeconomic backgrounds are both disproportionately likely to be inactive [@rowe2000sport; @farrell2014] and be affected by physical inactivity-related diseases [@petrovic2018contribution].
 Increasing the physical activity levels of the population is therefore high on the public health agenda: it not only has the potential to improve quality of life, reduce mortality rates and alleviate the strain on health and social care services, but also reduce the gap in health inequalities [@vilhelmsson2018].
 
-However, designing effective public health interventions that increase population physical activity is considerable challenge.[@reis2016] Implementing such interventions in a way that does not increase health inequalities might even be more difficult.
-Studies have shown that programmes to increase physical activity often fail to reach deprived communities and those most in need, suggesting access to physical activity facilities shows spatial and social inequality.[@white2009; @rideout2018] In this regard, parkrun, an international movement which organises free weekly 5 km running and walking events in public spaces, might provide valuable lessons.
+However, designing effective public health interventions that increase population physical activity is a considerable challenge [@reis2016; @kelly].
+Implementing such interventions in a way that does not increase health inequalities might even be more difficult.
+Studies have shown that programmes to increase physical activity often fail to reach deprived communities and those most in need, suggesting access to physical activity facilities shows spatial and social inequality [@white2009; @rideout2018].
+In this regard, parkrun, an international movement which organises free weekly 5 km running and walking events in public spaces, might provide valuable lessons.
 
-Since its founding in 2004, as a small event in London with 13 participants, parkrun has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@haake2018pr; @reece2019] The volunteer-led events are often characterised as accessible and inclusive.[@stevinson2013; @quirk2019] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2019changes; @wiltshire2018]
+Since its founding in 2004, as a small event in London with 13 participants, parkrun has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries [@haake2018pr; @reece2019].
+The volunteer-led events are often characterised as accessible and inclusive [@stevinson2013; @quirk2019].
+The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive [@stevinson2019changes; @wiltshire2018].
 
@@
 It might therefore be the case that parkrun events are primarily located in areas that are less deprived, while people living in more deprived communities may not have the same opportunities to participate.
-In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@reuters2018]
+In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socioeconomic groups [@reuters2018].
 
-The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England is equitable across areas with different levels of deprivation.
-Secondly, to identify 200 optimal locations for future events in order to improve geographic access, in particular for deprived communities.
+The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England is equitable across areas with different levels of deprivation; and secondly, to identify 200 optimal locations for future events in order to improve geographic access, in particular for deprived communities.
 
@@
 
-This study is a cross-sectional ecological analysis of the socio-economic disparities in geographic access to parkrun events in England at the end of 2018.
+This study is a cross-sectional ecological analysis of the socioeconomic disparities in geographic access to parkrun events in England at the end of 2018.
 All analyses were conducted on the level of Lower Layer Super Output Areas (LSOAs), which divide England into 32,844 geographic units which, on average, have a population of approximately 1,700.
-We assessed the relationship between access, defined as the distance (as the crow flies) to the nearest parkrun event, and socio-economic deprivation, measured using the Index of Multiple Deprivation.
+We assessed the relationship between access, defined as the distance (as the crow flies) to the nearest parkrun event, and socioeconomic deprivation, measured using the Index of Multiple Deprivation.
 In addition, we used information on public green spaces in England to conduct a simple location-allocation analysis, in order to identify 200 locations for future parkrun events that maximise access for the population.
@@
 
-  1) For all 32,844 LSOAs, we retrieved geographic locations, defined by the coordinates of its population-weighted centroid, 2017 total population estimates, and the 2015 Index of Multiple Deprivation (IMD) from the Office for National Statistics.[@ons2011; @ons2015; ons2017]
+  1) For all 32,844 LSOAs, we retrieved geographic locations, defined by the coordinates of its population-weighted centroid; 2017 total population estimates; and the 2015 Index of Multiple Deprivation (IMD) from the Office for National Statistics [@ons2011; @ons2015; @ons2017].
 
-  2) We included all 465 public parkrun events which were in operation in England by December 12th, 2018 - on this date, Sport England announced their plan to provide funding for setting up 200 additional parkrun events across England.[@parkrunGroup] The locations of the events were obtained from the parkrun UK website.[@parkrunweb]
+  2) We included all 465 public parkrun events which were in operation in England by December 12th, 2018 - on this date, Sport England announced their plan to provide funding for setting up 200 additional parkrun events across England [@parkrunGroup].
+The locations of the events were obtained from the parkrun UK website [@parkrunweb].
 
-  3) The locations of public green spaces in England were retrieved from an open dataset of Ordnance Survey.[@ordnance] parkrun events are held in various settings and terrains, and do not always require a single 5 km loop - some events have courses that involve running a combination of shorter loops..
+  3) The locations of public green spaces in England were retrieved from an open dataset of Ordnance Survey [@ordnance].
+parkrun events are held in various settings and terrains, and do not always require a single 5 km loop - some events have courses that involve running a combination of shorter loops.
 After evaluating existing parkrun event courses, we decided to consider all public parks, gardens, and playing fields in England with an area of 0.1 km$^2$ or more potentially suitable for hosting events (n=2,842).
@@
 
-  The socio-economic deprivation of LSOAs was measured using the 2015 IMD.
+  The socioeconomic deprivation of LSOAs was measured using the 2015 IMD.
 It is a measure of relative deprivation, which has been used in many similar studies.
 The IMD combines 37 indicators from seven domains (income, employment, education and skills, health and disability, crime, housing and services, and living environment) into a single score.
-The score ranges from 0 (least deprived) to 100 (most deprived).[@smith2015]
+The score ranges from 0 (least deprived) to 100 (most deprived) [@smith2015].
 
@@
 
-  More formally, the first step of the algorithm is defined by the following equation:
+  More formally, the first step of the algorithm evaluates the following expression:
 
@@
 
-  All data and the R source code that were used to generate the results of this study are provided on an open repository.[@githubrepo]
+  All data and the R source code that were used to generate the results of this study are provided on an open repository [@githubrepo].
 
@@
   Ethical approval was obtained from the Sheffield Hallam University Ethics Committee (ER10776545).
-We did not collect any personal information, but only used aggregate secondary data.
+We did not collect any personal information, and only used aggregate secondary data.
 The parkrun Research Board approved this research project, and four of its members (AMB, HQ, EG, SSJH) were actively involved in the interpretation of findings and writing of this manuscript.
@@
 
-  As of 12th December 2018, approximately 7%, 69%, and 91% of the English population lived within 1, 5, and 10 km distance of a parkrun event.
+  As of 12th December 2018, approximately 7%, 69%, and 91% of the English population lived within 1, 5, and 10 km distance of a parkrun event respectively.
 Only 578,043 people (1% of the English population) lived more than 20 km from an event.
-The average (SD) and median (IQR) distance to the nearest parkrun event were 4.65 (4.22) and 3.39 (1.99-5.83).
+The mean (SD) and median (IQR) distance to the nearest parkrun event were 4.65 (4.22) and 3.39 (1.99-5.83).
 The largest distance was observed for the 2,259 people living on the Isles of Scilly, who live about 76 km away from the next parkrun event on the mainland.
@@
 ![](./output/map_static.jpeg)
-**Figure 1**: Map of England showing current parkrun events (blue circles) and recommended new event locations (red triangles) ranked in descending order of estimated effect on overall population-weighted access.
-Information on all 200 identified optimal green space locations are provided in the appendix.
+**Figure 1: Map of England showing current parkrun events (blue circles) and recommended new event locations (red triangles) ranked in descending order of estimated effect on overall population-weighted access.
+Information on all 200 identified optimal green space locations are provided in the appendix.**
 
@@
 parkrun events are set up by volunteers, based on demand, not need.
-Studies have shown that the level of physical activity, and the availability of physical activity facilities generally declines with the level of deprivation.[@farrell2014socioeconomic] Opportunities for physical activity are often lacking in areas in most need.[e.g.
-@hillsdon2007] parkrun events, in contrast, seem to be often held in or near deprived areas, and are free to attend, giving anyone equal access, irrespective of their socio-economic background.
-Nevertheless, in a previous analysis, we found that participation in parkrun has a strong socioeconomic gradient with considerably higher participation rates in less deprived areas: about a third of all participants came from LSOAs in the least deprived quintile, while only 7% came from the most deprived quintile.[@smithschneider; @schneidersmith] This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been sufficient to engage people living in deprived communities.[@white2009] This means, creating additional events in optimal locations could improve overall geographic access further, yet effective strategies will still be needed to increase equity in engagement in new and current events.
+Studies have shown that the level of physical activity, and the availability of physical activity facilities generally declines with the level of deprivation [@farrell2014socioeconomic] Opportunities for physical activity are often lacking in areas in most need [e.g.
+@hillsdon2007].
+parkrun events, in contrast, seem to be often held in or near deprived areas, and are free to attend, giving anyone equal access, irrespective of their socioeconomic background.
+Nevertheless, in a previous analysis, we found that participation in parkrun has a strong socioeconomic gradient with considerably higher participation rates in less deprived areas: about a third of all participants came from LSOAs in the least deprived quintile, while only 7% came from the most deprived quintile [@smithschneider; @schneidersmith].
+This suggests that providing the opportunity to participate in parkrun events, while a necessary first step to enable participation, has not been sufficient to engage people living in deprived communities [@white2009].
+This means, creating additional events in optimal locations could improve overall geographic access further, yet effective strategies will still be needed to increase equity in engagement in new and current events.
 
@@
 Secondly the analysis makes use of large and rich datasets, with over 30,000 LSOA and over 400 existing parkrun events it is unlikely that individual outliers are affecting the results.
-The almost universal availability of parkrun events throughout the country provides a learning opportunity to explore socio-spatial determinants influencing physical activity behaviour on a national scale.[@wiltshire2018] Our study contributes to the limited research in this area and identifies possible leads for further investigation.
+The almost universal availability of parkrun events throughout the country provides a learning opportunity to explore socio-spatial determinants influencing physical activity behaviour on a national scale [@wiltshire2018].
+Our study contributes to the limited research in this area and identifies possible leads for further investigation.
 
@@
 Most importantly, geographic access is not measured as travel distance, or travel time, but as geodesic distance.
-In some cases, for example where natural barriers like hills, lakes or rivers block routes, the actual distance travelled may be far in excess of the geodesic distance.[@boscoe2012]
+In some cases, for example where natural barriers like hills, lakes or rivers block routes, the actual distance travelled may be far in excess of the geodesic distance [@boscoe2012].
 
@@
 
-  Finally.
-our analysis has been concerned only with determining to what extent deprived communities have geographic access to parkrun events.
+  Finally, our analysis has been concerned only with determining to what extent deprived communities have geographic access to parkrun events.
 We did not investigate what other factors independently explain access more generally.
@@
 
+<!-- UNBLINDED:
 ## Additional Information
@@
 Elizabeth Goyder; Roles: Supervision, Conceptualization, Writing – Review & Editing
+-->
 
@@
   kable_styling(font_size = 7,full_width = F) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(2:4, width = "1cm") %>%
-  column_spec(5:6, width = "1cm") 
+  column_spec(1, width = "4cm") %>%
+  column_spec(2:4, width = "1.5cm") %>%
+  column_spec(5:6, width = "1.5cm") 
 ```

### 2020-05-29 src/src4_analysis.R

@@ -73,6 +73,7 @@
     # as is
     fit1 = lm(mn_dstn ~ imd_sc,lsoa_sp,weights = pop)
     summary(fit1)
+    
     weightedCorr(lsoa_sp$imd_sc, lsoa_sp$mn_dstn,method = "pearson",weights = lsoa_sp$pop)
     weightedCorr(lsoa_sp$imd_sc, lsoa_sp$mn_dstn,method = "spearman",weights = lsoa_sp$pop)
     
@@ -95,23 +96,23 @@
 
 # 4  Density plot  ---------------
     # 99 rows removed
-    before_after_dist = ggplot(lsoa_sp@data) +
-      geom_density(aes(mn_dstn_new,fill="After 200 new parkrun events are set up"),col="black",alpha=0.5) +
-      geom_density(aes(mn_dstn,fill="Current access to parkrun events (December 12th, 2018)"),col="black",alpha=0.3) +
-      scale_fill_manual(values = c( "Current access to parkrun events (December 12th, 2018)" = "red","After 200 new parkrun events are set up"="lightgreen"),name="") +
-      scale_x_continuous(label=scales::number_format(accuracy = 1,suffix = " km"),
-                         limits = c(0,25),
-                         name = "Distance to the nearest parkrun event") +
-      theme_minimal()+
-      theme(legend.position = "bottom")
-
-      ggsave(plot=before_after_dist,filename = "./output/hist.png",height=6,width=8)
+    # before_after_dist = ggplot(lsoa_sp@data) +
+    #   geom_density(aes(mn_dstn_new,fill="After 200 new parkrun events are set up"),col="black",alpha=0.5) +
+    #   geom_density(aes(mn_dstn,fill="Current access to parkrun events (December 12th, 2018)"),col="black",alpha=0.3) +
+    #   scale_fill_manual(values = c( "Current access to parkrun events (December 12th, 2018)" = "red","After 200 new parkrun events are set up"="lightgreen"),name="") +
+    #   scale_x_continuous(label=scales::number_format(accuracy = 1,suffix = " km"),
+    #                      limits = c(0,25),
+    #                      name = "Distance to the nearest parkrun event") +
+    #   theme_minimal()+
+    #   theme(legend.position = "bottom")
+    # 
[…]
